---
title: "R is Your Friend Course Materials"
tbl-colwidths: [5,5,45,10,10,10,10,10]
---

## {{< fa calendar title="An calendar" >}} Schedule

This page contains an easy reference the topics, content, and assignments for a
semester.

{{< fa book >}} **Prepare** – readings and videos to complete before class

{{< fa laptop >}} **Slides** - presentation materials for class discussion / 
warm-up before the activity / lab

{{< fa clipboard >}} **Handout** - materials to accompany class discussion / 
warm-up before the activity / lab

{{< fa users >}} **Activity** – activities done during class

{{< fa pencil >}} **Lab** – end of week assignments

<!-- {{< fa pencil >}} **Exams** – end of unit practice exams -->

<!--
To update the calendar, add the relative path to the material's .qmd file
-->

```{r}
#| label: packages
#| echo: false
#| message: false
#| warning: false

library(gt)
library(tidyverse)
```


```{r}
#| label: read-data
#| echo: false

d <- readxl::read_excel("index.xlsx")

```

```{r}
#| label: format-calendar-entries
#| echo: false

t <- d |>
  select(-`practice exam`) |> 
  mutate(
    week = str_c("Week ", week),
    day = str_c("Day ", day),
    week_label = case_when(week == "Week 1" & day == "Day 1" ~ week, 
                           week != lag(week) ~ week,
                           .default = ""), 
    unit = recode_values(unit, 
                         1 ~ "Unit 1", 
                         2 ~ "Unit 2", 
                         3 ~ "Unit 3"
                         ), 
    unit = fct_relevel(unit,
                     "Unit 1", 
                     "Unit 2",
                     "Unit 3"
                     )
  ) |> 
  select(unit, 
         week_label, 
         day, 
         topic, 
         preparation, 
         slides, 
         activity, 
         lab, 
         handout) |> 
  gt() |> 
  cols_width(
    week_label:day ~ pct(3),
    topic ~ pct(15),
    preparation:handout ~ pct(6)
  ) |> 
  text_transform(cells_body(columns = c(preparation)),
                 fn = function(x){
                   ifelse(is.na(x), fontawesome::fa('book', fill_opacity = 0.1), 
                          glue::glue('<a href="{x}"><i class="fa fa-book" style="color: purple;"></i></a>'))
                 }) |> 
  text_transform(cells_body(columns = c(slides)),
                 fn = function(x){
                   ifelse(is.na(x), fontawesome::fa('laptop', fill_opacity = 0.1), 
                           glue::glue('<a href="{x}"><i class="fa fa-laptop" style="color: purple;"></i></a>'))
                 }) |> 
  text_transform(cells_body(columns = c(activity)),
                 fn = function(x){
                   ifelse(is.na(x), fontawesome::fa('users', fill_opacity = 0.1),
                          glue::glue('<a href="{x}"><i class="fa fa-users" style="color: purple;"></i></a>'))
                  }) |> 
  text_transform(cells_body(columns = c(lab)),
                 fn = function(x){
                   ifelse(is.na(x), fontawesome::fa('pencil', fill_opacity = 0.1), 
                          glue::glue('<a href="{x}"><i class="fa fa-pencil" style="color: purple;"></i></a>'))
                 }) |> 
  text_transform(cells_body(columns = c(handout)),
                 fn = function(x){
                   ifelse(is.na(x), fontawesome::fa('clipboard', fill_opacity = 0.1), 
                          glue::glue('<a href="{x}"><i class="fa fa-clipboard" style="color: purple;"></i></a>'))
                 }) |> 
  sub_missing(columns = topic, missing_text = "") |> 
  cols_align(align = "center",
             columns = !topic) |> 
  tab_row_group(
  label = "Unit 1",
  rows = (unit == "Unit 1"), 
  id = "Unit 1"
  ) |> 
  tab_row_group(
  label = "Unit 2",
  rows = (unit == "Unit 2"), 
  id = "Unit 2"
  ) |> 
  tab_row_group(
  label = "Unit 3",
  rows = (unit == "Unit 3"), 
  id = "Unit 3"
  ) |> 
  cols_hide(unit) |> 
  tab_style(
    style = list(
    cell_fill(color = "gray95"),
    cell_text(size = "x-large")
    ),
    locations = cells_row_groups(groups = c("Unit 1",
                                            "Unit 2",
                                            "Unit 3"
                                            )
                                 )
    ) |>
  row_group_order(groups = c("Unit 1",
                             "Unit 2",
                             "Unit 3")
                  ) |> 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = topic)
    ) |> 
   tab_style(
     style = list(
      cell_fill(color = "#E8EEF6")
      ),
    locations = cells_body(
      rows = str_detect(topic, pattern = "^Project [0-9]+ Due$")
    )
  ) |> 
  tab_options(
    row_group.padding = px(8)
  ) |> 
  cols_label(
    week_label = "",
    topic = "",
    day = "",
    preparation = md("**Prepare**"),
    slides = md("**Slides**"), 
    handout = md("**Handout**"),
    activity = md("**Activity**"),
    lab = md("**Lab**")
    ) |> 
  as_raw_html() 
```

```{r}
#| echo: false
#| results: "asis"
#| label: output-calendar

cat(t)
```


