---
title: "Working with R Objects"
format: 
  revealjs:
      theme: [simple, style.scss]
      auto-stretch: false
editor: source
---

## Types of Objects in R

This week you've learned how to make four different types of objects in R:

- vectors
- matrices
- lists
- dataframes

Let's review each of these so they are fresh in our memory!

## Vectors

Suppose you want to create two vectors:

1. The days of the week
2. The temperature of the water at Pismo each day

Pencil out what code you would write to create these vectors. 

## Creating Vectors

::: columns
::: {.column width="48%"}
```{r}
#| label: create-vector-days
#| echo: true
#| code-line-numbers: false

days <- c("Monday", 
          "Tuesday", 
          "Wednesday", 
          "Thursday", 
          "Friday", 
          "Saturday", 
          "Sunday")
```
:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
```{r}
#| label: create-vector-temps
#| echo: true
#| code-line-numbers: false

temps <- c(60.1, 
           59.4, 
           60.8, 
           61, 
           60.1, 
           62.1, 
           59.7)
```
:::
:::

. . .

</br>

What would you expect this code to output?

::: columns
::: {.column width="48%"}
```{r}
#| label: check-data-type-days
#| echo: true
#| eval: false
#| code-line-numbers: false

typeof(days)
```
:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
```{r}
#| label: check-data-type-temps
#| echo: true
#| eval: false
#| code-line-numbers: false

typeof(temps)
```

:::
:::

## Matrices

A matrix is a two dimensional vector! Let's assess our understanding of how 
matrices are made. 

What would you expect this code to output?

```{r}
#| echo: true
#| eval: false
#| label: matrix-wrong-length
#| code-line-numbers: false

matrix(temps, nrow = 2)
```

. . .

</br>

```{r}
#| error: true
#| warning: true
#| echo: false

matrix(temps, nrow = 2)
```

## Lists

Unlike matrices, lists allow the elements to have different data types **and**
different sizes. 

</br>

Here are two ways to make a list:

::: columns
::: {.column width="48%"}
```{r}
#| label: create-unnamed-list
#| echo: true
#| code-line-numbers: false

watertemp_list <- list(days, 
                       temps)
```

:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
```{r}
#| label: create-named-list
#| echo: true
#| code-line-numbers: false

watertemp_list <- list(days = days, 
                       temps = temps)
```
:::
:::

</br>

**How are these lists different?**

## Dataframes

A dataframe is a special type of list---where every element (column) is the
same size. 

</br>

Draw out how the `watertemp_df` would look:

```{r}
#| label: make-dataframe
#| echo: true
#| code-line-numbers: false

watertemp_df <- data.frame(days = days, 
                           temps = temps)
```

## Assessing the Size of the Object

We might want to know how big an object is. 

What would you expect these lines of code to output?

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: false

length(days)
```

</br>

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: false

length(watertemp_list)
```

</br>

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: false

length(watertemp_df)
```

## Assessing the Size of a List or Dataframe

::: columns
::: {.column width="48%"}
**Lists**

```{r}
#| code-line-numbers: false

lengths(watertemp_list)
```

Notice we are using `length`[`s`]{.underline} here, not `length`!

:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
::: {.fragment}
**Dataframes**

```{r}
#| label: dimensions-of-df
#| code-line-numbers: false

dim(watertemp_df)

nrow(watertemp_df)

ncol(watertemp_df)
```
:::
:::
:::

## Accessing the Elements of an Object

We might want to extract components of an object. 

Pencil out how you would write code to do the following:

- Extract the first two days of the `days` vector.
- Extract the first element of the `watertemps_list` object.
- Extract the second row of the `watertemps_df` object.

## Accessing the Elements of an Object - Solution

- Extract the first two days of the `days` vector.

::: columns
::: {.column width="48%"}
```{r}
#| label: extract-vector-option1
#| code-line-numbers: false

days[1:2]

```

:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
```{r}
#| label: extract-vector-option2
#| code-line-numbers: false

days[c(1, 2)]

```
:::
:::

. . .

- Extract the first element of the `watertemps_list` object.

::: columns
::: {.column width="48%"}
```{r}
#| label: extract-list-option1
#| code-line-numbers: false

watertemp_list$days
```
:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
```{r}
#| label: extract-list-option2
#| code-line-numbers: false

watertemp_list[[1]]
```

:::
:::

## Accessing the Elements of an Object - Solution

- Extract the second row of the `watertemps_df` object.

```{r}
#| label: extract-df
#| code-line-numbers: false

watertemp_df[2, ]
```


## Modifying an Object

Suppose we wanted to add a column containing information on the air temperature 
to `watertemp_df`. 

</br>

Here are the air temperatures for each day: 67, 65, 67, 70, 69, 68, 68

</br>

Pencil out how you would add this information as a new column in `watertemp_df`.

## Modifying an Object -- Adding a Column

```{r}
#| echo: true
#| label: add-airtemp-column
#| code-line-numbers: false

watertemp_df$airtemp <- c(67, 65, 67, 70, 69, 68, 68)

watertemp_df
```

<!-- Allison: I thought this might be of interest, but also might be too much -->
<!-- ## Changing Column Names -->

<!-- Since the dataframe now has information on **two** types of temperatures, it -->
<!-- feels like our original column name (`temps`) is too vague.  -->

<!-- If we wanted to change the name of the columns, there are two options: -->

<!-- 1. Go back and change the name when we originally made `watertemp_df` -->

<!-- 2. Change the names now -->

<!-- Option 2 would look something like this: -->

<!-- ```{r} -->
<!-- #| code-line-numbers: false -->

<!-- names(watertemp_df) <- c("days", "watertemp", "airtemp") -->
<!-- watertemp_df -->
<!-- ``` -->

## Modifying an Object -- Changing Specific Values

We also learned that we got the water temperature wrong on Thursday and 
Friday. 

On Thursday the water temperature was actually 60.8 and on Friday the
temperature was actually 61.2. 

</br>

Pencil out how you would adjust these water temperature values. 

## Modifying an Object -- Changing Specific Values

- Thursday is row 4 and Friday is row 5
- water temperature is column 2 (named `temps`)

**Option 1:** 

```{r}
#| label: change-temps-option1
#| code-line-numbers: false

watertemp_df[4:5, 2] <- c(60.8, 61.2)
```

**Option 2:** 

```{r}
#| label: change-temps-option2
#| code-line-numbers: false

watertemp_df[4:5, "temps"] <- c(60.8, 61.2)
```

**Option 3:** 

```{r}
#| label: change-temps-option3
#| code-line-numbers: false

watertemp_df$temps[4:5] <- c(60.8, 61.2)
```


## Modifying an Object -- Changing Column Values

We originally store the air and water temperatures in degrees Fahrenheit, but
many people prefer degrees Celsius. 

The formula for converting from Fahrenheit to Celsius is:

$$\text{Celsius} = \frac{5}{9}(\text{Fahrenheit} - 32)$$

Pencil out how you would convert the `temps` and `airtemp` columns to Celsius.

## Modifying an Object -- Changing Column Values

**Option 1:** Adjust the temperatures and save them back in the original column

```{r}
#| label: changing-column-values
#| code-line-numbers: false

watertemp_df$temps <- (5/9) * (watertemp_df$temps  - 32)
watertemp_df$airtemp <- (5/9) * (watertemp_df$airtemp  - 32)
```

</br>

**Option 2:** Make a new column with the Celsius temps

```{r}
#| label: creating-new-column-values
#| code-line-numbers: false
#| eval: false

watertemp_df$temps_c <- (5/9) * (watertemp_df$temps  - 32)
watertemp_df$airtemp_c <- (5/9) * (watertemp_df$airtemp  - 32)
```

## Assessing Values of an Object

Personally, I won't swim at Pismo if the temperature is below 16 degrees 
(Celsius). How many days in the last week would I have swam?

Pencil out the code you would write to answer this question.

1. What relational comparison will you use?
2. What output will this comparison give you?
2. How will you use this output to find the number of days I would have swam?

## How many days would I have swam?

1. What relational comparison will you use?

```{r}
#| label: relational-comparison
#| code-line-numbers: false
#| eval: false

watertemp_df$temps >= 16
```

2. What output will this comparison give you?

```{r}
#| label: logical-values
#| code-line-numbers: false
#| echo: false

watertemp_df$temps >= 16
```

3. How will you use this output to find the number of days I would have swam?

```{r}
#| label: sum-logical-values
#| code-line-numbers: false

sum(watertemp_df$temps >= 16)
```
