---
title: "Control Flow"
format: 
  revealjs:
      theme: [simple, style.scss]
      auto-stretch: false
editor: source
---

## Warming Up for the Practice Activity

This week's practice activity combines **conditional statements** and 
**iteration**. Let's start by practicing setting up a couple of 

## Problem A - Fizz Buzz

Suppose you have a [vector]{.underline} of numbers (e.g., `1:100`). For each
value, you want to print the following output:

::: {.incremental}
- if a number is divisible by 3, print `"Fizz"`,
- if the number is divisible by 5, print `"Buzz"`,
- if the number is divisible by **both** 3 and 5, print `"FizzBuzz"`, 
- if none of these are true then print the number.
:::

. . .

::: {.callout-tip}
# Pencil it Out

Brainstorm how you would write R code to solve this problem. **Your solution 
must use a `for()`-loop**.
:::

## Problem B - Dog Sitting

Suppose you have a [dataframe]{.underline} with information about dogs you
are currently pet sitting. 

```{r}
#| label: create-dataframe
#| echo: false

dogs <- data.frame(
  dog = c("Annie", "Shilo", "Buddy", "Pepper", "Ruby", "Wallie", "Chompers"), 
  size = c("small", "medium", "large", "medium", "small", "small", "large"), 
  age = c(3, 11, 15, 5, 7, 10, 12)
)
```

```{r}
#| code-line-numbers: false

dogs
```

You need to know how much dog food you need to have this week to feed all these 
doggies!

## Problem B - Dog Sitting

Here are the food needs based on the dog's size: 

- Small dogs: 1 cup (8 ounces)
- Medium dogs: 2 cups (16 ounces)
- Large dogs: 4 cups (32 ounces)

If the dog is older than 12, they need to eat the same portions, but in wet food
(instead of dry food). 

</br>

::: {.callout-tip}
# Pencil it Out

Brainstorm how you would write R code to figure out how much total dog food 
(both wet and dry) you need to have on hand. **Your solution must use a `for()`-loop**.
:::

## Work Session 1 

When is your birthday?

- Before July 1? **Start with Problem A (FizzBuzz)**. 

- On or after July 1? **Start with Problem B (Dog Sitter)**. 

. . .

</br>

::: {.callout-tip}
# Start with a Gameplan!

Before you start writing code, write down the steps you need to take to 
accomplish the task!

- What conditions do you need to check?
- What should happen if the condition is met?
- What order should the conditions be checked?
:::

```{r}
#| echo: false
countdown::countdown(minutes = 15)
```

## Work Session 2

Work on the problem you were not initially assigned!

</br>

::: {.callout-tip}
# Start with a Gameplan!

Before you start writing code, write down the steps you need to take to 
accomplish the task!

- What conditions do you need to check?
- What should happen if the condition is met?
- What order should the conditions be checked?
:::

```{r}
#| echo: false
countdown::countdown(minutes = 15)
```

## Compare with a Peer

Pair up with **one** person and compare your solution strategies for Problem A
and Problem B. 

```{r}
#| echo: false
countdown::countdown(minutes = 10)
```

## Fizz Buzz

::: columns
::: {.column width="75%"}
```{r}
#| code-line-numbers: false
#| eval: false
#| label: code-fizzbuzz

num <- 1:100

for(i in 1:length(num)){
  # Numbers divisible by both 3 and 5
  if(num[i] %% 3 == 0 & num[i] %% 5 == 0){
    print("FizzBuzz")
    } 
  # Numbers divisible by 5
  else if(num[i] %% 5 == 0){
    print("Buzz")
    }
  # Numbers divisible by 3
  else if(num[i] %% 3 == 0){
    print("Fizz")
    } 
  # Numbers not divisible by 3 or 5
  else{
    print(num[i])
    }
  }

```

:::

::: {.column width="5%"}
:::

::: {.column width="20%"}
::: {.fragment}
```{r}
#| echo: false
#| label: output-fizzbuzz

num <- 1:100

for(i in 1:length(num)){
  # Numbers divisible by both 3 and 5
  if(num[i] %% 3 == 0 & num[i] %% 5 == 0){
    print("FizzBuzz")
    } 
  # Numbers divisible by 5
  else if(num[i] %% 5 == 0){
    print("Buzz")
    }
  # Numbers divisible by 3
  else if(num[i] %% 3 == 0){
    print("Fizz")
    } 
  # Numbers not divisible by 3 or 5
  else{
    print(num[i])
    }
  }

```
:::
:::
:::

## Dog Sitter

```{r}
#| code-line-numbers: false
#| label: code-dog-food

total_wetfood <- 0
total_dryfood <- 0

for(i in 1:nrow(dogs)){
  # Small dogs
  if(dogs$size[i] == "small"){
    # Old dogs
    if(dogs$age[i] > 12){
      total_wetfood <- total_wetfood + 8
    } 
    # Younger dogs
    else{
      total_dryfood <- total_dryfood + 8
    }
  } 
  # Medium dogs
  else if(dogs$size[i] == "medium"){
    # Older dogs
    if(dogs$age[i] > 12){
      total_wetfood <- total_wetfood + 16
    } 
    # Younger dogs
    else{
      total_dryfood <- total_dryfood + 16
      }
  } 
  # Large dogs
  else{
    # Old dogs
    if(dogs$age[i] > 12){
      total_wetfood <- total_wetfood + 32
    } 
    # Younger dogs
    else{
      total_dryfood <- total_dryfood + 32
      }
    }
}
```

. . .

::: columns
::: {.column width="48%"}
```{r}
#| code-line-numbers: false

total_wetfood
```

:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
```{r}
#| code-line-numbers: false

total_dryfood
```

:::
:::


## How would I...

use a `while()`-loop to figure out the first dog that needed wet food?

. . .

</br>

::: columns
::: {.column width="48%"}
**Initialize Values**
```{r}
#| code-line-numbers: false
#| label: while-loop-initialize

age <- ""
counter <- 0
```

::: {.fragment}
**Create the Loop**
```{r}
#| code-line-numbers: false
#| label: while-loop

while(age < 12){
  counter <- counter + 1
  age <- dogs$age[counter]  
}
```
:::
:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
::: {.fragment}
**See the Result**
```{r}
#| label: while-loop-result
#| code-line-numbers: false

dogs[counter,]
```
:::
:::
:::

## Dataframe Review

Suppose we have a dataframe about cars:

```{r}
#| label: mpg-data
#| code-line-numbers: false
#| results: hide

library(ggplot2)
mpg
```

```{r}
#| label: pretty-df
#| echo: false

mpg |> 
  dplyr::slice_sample(n = 12) |> 
  gt::gt() |> 
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_labels(columns = everything())
    )
```

## Grabbing Information

What code would I run to learn the number of rows the `mpg` dataframe has?
What about the number of columns?

. . .

::: columns
::: {.column width="30%"}
```{r}
#| code-line-numbers: false

dim(mpg)
```

:::

::: {.column width="3%"}
:::

::: {.column width="30%"}
```{r}
#| code-line-numbers: false

nrow(mpg)
```
:::

::: {.column width="3%"}
:::

::: {.column width="30%"}
```{r}
#| code-line-numbers: false

ncol(mpg)
```
:::
:::

. . .

What code would I write to grab the `model` column from the `mpg` dataframe?

. . .

::: columns
::: {.column width="48%"}
```{r}
#| code-line-numbers: false
#| eval: false

mpg$model
```

:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
```{r}
#| code-line-numbers: false
#| eval: false

mpg[["model"]]
```
:::
:::

. . .

How would I modify this code to grab the model of the fourth car in the `mpg`
dataframe?

. . .

::: columns
::: {.column width="48%"}
```{r}
#| code-line-numbers: false
#| eval: false

mpg$model[4]
```

:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
```{r}
#| code-line-numbers: false
#| eval: false

mpg[["model"]][4]
```
:::
:::

