---
title: "Week 2: Working with R Objects"
format: html
execute: 
  eval: false
---

## Vectors

Pencil out what code you would write to create two vectors:

1. The days of the week
2. The temperature of the water at Pismo each day


</br>
</br>
</br>

## Matrices

What would you expect this code to output?

```{r}
#| echo: true
#| eval: false
#| label: matrix-wrong-length

matrix(temps, nrow = 2)
```


## Lists

Here are two ways to make a list:

::: columns
::: {.column width="48%"}
```{r}
#| label: create-unnamed-list
#| echo: true
#| code-line-numbers: false
#| eval: false

list(days, temps)
```

:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
```{r}
#| label: create-named-list
#| echo: true
#| code-line-numbers: false
#| eval: false

list(days = days, 
     temps = temps)
```
:::
:::

</br>

**How are these lists different?**

</br>
</br>


## Dataframes

Draw out what this code would output:

```{r}
#| label: make-dataframe
#| eval: false
#| echo: true
#| code-line-numbers: false

data.frame(days = days, 
           temps = temps)
```

</br>
</br>
</br>

## Assessing the Size of the Object

What would you expect these lines of code to output?

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: false

length(days)
```

</br>

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: false

length(watertemp_list)
```

</br>

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: false

length(watertemp_df)
```

## Assessing the Size of a List or Dataframe

What are the functions that help you access information about a list or 
dataframe?

</br>
</br>

## Accessing the Elements of an Object

We might want to extract components of an object. 

Pencil out how you would write code to do the following:

- Extract the first two days of the `days` vector.

</br>

- Extract the first element of the `watertemps_list` object.

</br>

- Extract the second row of the `watertemps_df` object.

</br>

## Modifying an Object -- Adding a Column

Suppose we wanted to add a column containing information on the airtemperature 
to `watertemp_df`. 

Here are the air temperatures for each day: 67, 65, 67, 70, 69, 68, 68

Pencil out how you would add this information as a new column in `watertemp_df`.

</br>
</br>

## Modifying an Object -- Changing Specific Values

We also learned that we got the water temperature wrong on Thursday and 
Friday. 

On Thursday the water temperature was actually 60.8 and on Friday the
temperature was actually 61.2. 

Pencil out how you would adjust these water temperature values in the 
`watertemp_df` dataframe.

</br>
</br>

## Modifying an Object -- Changing Column Values

We originally store the air and water temperatures in degrees Fahrenheit, but
many people prefer degrees Celsius. 

The formula for converting from Fahrenheit to Celsius is:

$$\text{Celsius} = \frac{5}{9}(\text{Fahrenheit} - 32)$$

Pencil out how you would convert the `temps` and `airtemp` columns of the
`watertemp_df` dataframe to Celsius.

</br>
</br>
</br>
</br>

## Assessing Values of an Object

Personally, I won't swim at Pismo if the temperature is below 17 degrees 
(Celsius). How many days in the last week would I have swam?

Pencil out the code you would write to answer this question.

1. What relational comparison will you use?
2. What output will this comparison give you?
2. How will you use this output to find the number of days I would have swam?

</br>
</br>
