---
title: "Week 3 Part 1: Control Flow"
format: html
---

This week's practice activity combines **conditional statements** and 
**iteration**. Let's start by practicing setting up a couple of 

## Problem A - Fizz Buzz

Suppose you have a [vector]{.underline} of numbers (e.g., `1:100`). For each
value, you want to print the following output:

- if a number is divisible by 3, print `"Fizz"`,
- if the number is divisible by 5, print `"Buzz"`,
- if the number is divisible by **both** 3 and 5, print `"FizzBuzz"`, 
- if none of these are true then print the number.

## Problem B - Dog Sitting

Suppose you have a [dataframe]{.underline} with information about dogs you
are currently pet sitting. 

```{r}
#| label: create-dataframe
#| echo: false

dogs <- data.frame(
  dog = c("Annie", "Shilo", "Buddy", "Pepper", "Ruby", "Wallie", "Chompers"), 
  size = c("small", "medium", "large", "medium", "small", "small", "large"), 
  age = c(3, 11, 15, 5, 7, 10, 12)
)
```

```{r}
#| label: dog-preview

dogs
```

You need to know how much dog food you need to have this week to feed all these 
doggies! Here are the food needs based on the dog's size: 

- Small dogs: 1 cup (8 ounces)
- Medium dogs: 2 cups (16 ounces)
- Large dogs: 4 cups (32 ounces)

If the dog is older than 12, they need to eat the same portions, but in wet food
(instead of dry food). 


## Work Session 1 

</br>
</br>
</br>
</br>
</br>
</br>
</br>
</br>
</br>

## Work Session 2

</br>
</br>
</br>
</br>
</br>
</br>
</br>
</br>
</br>

## How would I...

use a `while()`-loop to figure out the first dog that needed wet food?

</br>
</br>
</br>
</br>
</br>
</br>


## Dataframe Review

Suppose we have a dataframe about cars:

```{r}
#| label: mpg-data
#| code-line-numbers: false
#| results: hide

library(ggplot2)
mpg
```

```{r}
#| label: pretty-df
#| echo: false

mpg |> 
  dplyr::slice_sample(n = 12) |> 
  gt::gt() |> 
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_labels(columns = everything())
    )
```

## Grabbing Information

What code would I run to learn the number of rows the `mpg` dataframe has?
What about the number of columns?

</br>
</br>

What code would I write to grab the `model` column from the `mpg` dataframe?

</br>
</br>

How would I modify this code to grab the model of the fourth car in the `mpg`
dataframe?

</br>
</br>
