---
title: "Handout 7: More Data Wrangling"
format: html
---

```{r}
#| label: setup
#| echo: false

library(tidyverse)
library(liver)
library(knitr)
data(cereal)
```

:::callout-warning

In progress! For class, I am imagining this would be rendered as a .docx and need some editing to be filled out by students.

Also considering -- too long for finishing PA7 as well? Maybe students will need to finish this PA outside of class or on the second day of class this week?
:::

## Extending dplyr verbs

1.  What is this code trying to do and why doesn't it work? How can we fix the code?

```{r}
cereal |> 
  select(sugars) |> 
  sd(na.rm = T)
```

2.  When would I want to use `count()` versus `summarize(n = n())`?

3.  How do columns need to be input into the `across()` function?

4.  How do functions need to be input into the `across()` function?

5.  What is the general structure of a lambda function? What does the `x` inside the lambda function represent?

Remember, you got warnings in practice activity 6 when converting some columns to numeric? If you look at the original data, you can see this is because missing values were indicated with the string `"NULL"`.

We could drop these rows before converting the columns to numeric if desired, using `if_any()`:

```{r}
#| echo: true
#| eval: false

colleges_clean <- colleges |> 
  filter(
    !if_any(.cols = ADM_RATE:TUITIONFEE_OUT, 
            .fns = \(x) x == "NULL")
    ) 
```

6.  How would you describe what `\(x) x == "NULL"` is doing?

7.  What does the `!` in front of the `if_any()` do?

## Reshaping Data

The slides contain the first couple of rows of the cereal data.

1.  Is the cereal data wide or long?

I want to create the following plot:

```{r}
#| echo: false

cereal |> 
  pivot_longer(cols = calories:vitamins,
               names_to = "Nutrient",
               values_to = "Amount") |> 
  group_by(shelf, Nutrient) |> 
  summarise(mean_amount = mean(Amount)) |> 
  ggplot(aes(x = shelf, 
             y = mean_amount, 
             color = Nutrient)) +
  geom_point() +
  geom_line() +
  labs(x = "Shelf", y = "", subtitle = "Mean Amount")
```

2.  With the original `cereal` data, how would I create this plot? What geometries and aesthetics would I use? You can write a couple of lines of code or sketch a game plan.

3.  Sketch the data that you would **want** to create this plot. What columns do you want and what rows?

4.  Complete the following code to reshape the cereal data for the plot:

```{r}
#| eval: false

cereal |> 
  pivot_      (cols =                        ,
               names_to =              
               values_to =                  )
```

5.  Here are the first couple of rows of the cereal data with relevant variables:

<!--I don't want students to get caught up in the extra variables -- weight, cups, and rating, but is this confusing?-->

```{r}
#| echo: false

cereal |> 
  relocate(shelf, .after = type) |> 
  select(-weight, -cups, -rating) |> 
  slice_head(n = 3) |> 
  kable()
```

Fill in the table below with a couple of rows of the data if we reshape it using the code above!

| name | manuf | type | shelf | ______  | _____ |
|------|------|------|------|------|------|
|      |      |      |      |      |      |
|      |      |      |      |      |      |
|      |      |      |      |      |      |
|      |      |      |      |      |      |
|      |      |      |      |      |      |
|      |      |      |      |      |      |
|      |      |      |      |      |      |
|      |      |      |      |      |      |
|      |      |      |      |      |      |
|      |      |      |      |      |      |
