---
title: "Handout 7: More Data Wrangling"
format: html
---


```{r}
#| label: setup

library(tidyverse)
library(liver)
data(cereal)
```

## Extending dplyr verbs

1. What is this code trying to do and why doesn't it work? How can we fix the code?

```{r}
cereal |> 
  select(sugars) |> 
  sd(na.rm = T)
```

2. SOMETHING ABOUT count

3.  How do columns need to be input into the `across()` function?

4. How do functions need to be input into the `across()` function?

5. What does the `x` inside the lambda function represent?

Remember, you got warnings in practice activity 6 when converting some columns to numeric? If you look at the original data, you can see this is because missing values were indicated with the string `"NULL"`.

We could drop these rows before converting the columns to numeric if desired, using `if_any()`:

```{r}
#| echo: true
#| eval: false

colleges_clean <- colleges |> 
  filter(
    !if_any(.cols = ADM_RATE:TUITIONFEE_OUT, 
            .fns = \(x) x == "NULL")
    ) 
```

6. How would you describe what `\(x) x == "NULL"` is doing?

7. What does the `!` in front of the `if_any()` do?


## Reshaping Data

The slides contain the first couple of rows of the cereal data. 

1. Is the cereal data wide or long?


I want to create the following plot: 

```{r}
#| echo: false

cereal |> 
  pivot_longer(cols = calories:vitamins,
               names_to = "Nutrient",
               values_to = "Amount") |> 
  group_by(shelf, Nutrient) |> 
  summarise(mean_amount = mean(Amount)) |> 
  ggplot(aes(x = shelf, 
             y = mean_amount, 
             color = Nutrient)) +
  geom_point() +
  geom_line() +
  labs(x = "Shelf", y = "", subtitle = "Mean Amount")
```



2. With the original `cereal` data, how would I create this plot? What geometries and aesthetics would I use? You can write a couple of lines of code or sketch a game plan.


3. Sketch the data that you would **want** to create this plot. What columns do you want and what rows?


4. manually draw out converting data from wide to long

5. Complete the following code to reshape the cereal data:

```{r}
#| eval: false

cereal |> 
  pivot_      (cols =                        ,
               names_to =              
               values_to =                  )
```

