---
title: "Code-Along 8: Factor Variables"
format: 
  html:
    code-tools: true
    toc: true      
execute: 
  error: true
  echo: true
  message: false
  warning: false

author: Your Name
embed-resources: true
---
```{r}
#| label: setup

library(tidyverse)
library(readxl)

eras_data <- read_xlsx("../../data/08-TS_data.xlsx")
```

<!--
Note: edit path above to read straight from GitHub?
-->

Let's explore factor variables in R! Work through these questions, running or editing the code provided to help you answer them.

1. What are the two general purposes for factor variables (in R)?

### `fct()`

2. The `fct()` function in **forcats** serves the same purpose as what function in base R?

3. How are the levels automatically ordered with the `fct()` function? How is this different from base R?

```{r}
eras_data |> 
  mutate(Album = fct(Album)) |> 
  pull(Album)
```

4. Edit the code above to just output the **levels** of the `Album` variable.

5. What does including the `levels = ` argument do in `fct()`?

```{r}
eras_data |> 
  mutate(Album = fct(Album,
                     levels = c("Fearless","Speak Now","Red",
                                "1989", "Reputation","Lover",
                                "Folklore", "Evermore","Midnights"))) |> 
  pull(Album)
```

5. What happens if you forget to provide an album to `levels = ` that exists in the data in the code above?

### `fct_recode()`

6. What does this function do? When would you want to use it?

7. What happens to non-specified levels in `fct_recode()`?

8. Change the level "Red" to "Red!" and "Fearless" to "Fearless - Taylors Version" in the `Album` variable:

```{r}
```

### `fct_collapse()`

9. What does this function do?  When would you want to use it?

```{r}
eras_data |> 
  mutate(Genre = fct_collapse(.f= Album,
                       "country pop" = c("Fearless"),
                       "pop rock" = c("Speak Now","Red"),
                       "electropop" = c("1989","Reputation","Lover"),
                       "folk pop" = c("Folklore","Evermore"),
                       "alt-pop" = "Midnights"))
```

10. Why would you want to use `fct_collapse()` rather than `case_when()`? 

11. What happens to levels of the variable `.f = ` that you do not include in `fct_collapse()`?
 
```{r}
eras_data |> 
  mutate(genre = fct_collapse(.f= Album,
                       "country pop" = c( "Fearless"),
                       "pop rock" = c("Speak Now","Red"),
                       "electropop" = c("1989","Reputation","Lover"),
                       "folk pop" = c("Folklore","Evermore")
                       )) |> 
  pull(genre) |> 
  levels()
``` 
 
12. Edit the code below to fix this, so that the Midnights album is categorized as "other" in the `genre` variable.

```{r}
eras_data |> 
  mutate(genre = fct_collapse(.f= Album,
                       "country pop" = c( "Fearless"),
                       "pop rock" = c("Speak Now","Red"),
                       "electropop" = c("1989","Reputation","Lover"),
                       "folk pop" = c("Folklore","Evermore")
                       )) |> 
  pull(genre) |> 
  levels()
``` 
 
### `fct_relevel()` 

13. What does this function do?  When would you want to use it?

```{r}
eras_data |> 
  mutate(Album = fct_relevel(.f = Album, 
                             c("Fearless","1989","Taylor Swift",
                               "Speak Now","Red","Midnights","Reputation"))) |>
  pull(Album) |>
  levels()
```

14. What happens with levels that are not specified?

### `fct_infreq()`

15. What does `fct_infreq()` do?  When would you want to use it?

```{r}
eras_data |> 
  ggplot() +
  geom_bar(aes(y = fct_infreq(Album)), 
           fill = "#A5C9A5") +
  theme_minimal() +
  labs(x = "Number of Songs",
       y = "",
       subtitle = "Album",
       title = "Songs Played on the Eras Tour")
```

### `fct_reorder()`

16. What does `fct_reorder()` do?  In what setting would you use it?

```{r}
# you may need to install this package!
library(ggridges)

eras_data |> 
  ggplot(aes(x = Length, 
             y = fct_reorder(.f = Album,
                             .x = Length,
                             .fun = mean), 
             fill = Album)) +
  geom_density_ridges() +
  theme_minimal() +
  theme(legend.position = "none")+
  labs(x = "Song Length (mins)",
       y = "",
       subtitle = "Album",
       title = "Songs Played on the Eras Tour")
```

### `fct_reorder2()`

17. What does `fct_reorder2()` do?  In what setting would you use it?

18. What are the arguments to `fct_reorder2()`?

19. Let's improve a plot of the number of listings over time for five cities in Texas. Use `fct_reorder2()` to reorder the legend below so that the legend aligns with the order of the lines in the plot.

```{r}
data(txhousing)

txhousing |>
  filter(city %in% c("Dallas","Fort Worth", "Austin",
                     "Houston", "El Paso"),
         !is.na(listings)) |> 
  ggplot() +
  geom_line(mapping = aes(x = date, 
                          y = listings,
                          color = city))
```
