---
title: "PA 9: Cancel Culture"
subtitle: "Strings and Date Variables"
format: 
  html:
    code-tools: true
    toc: true      
execute: 
  error: true
  echo: true
  message: false
  warning: false

author: Your Name
embed-resources: true
---
```{r setup}
# load packages
library(tidyverse)
```

The data are compiled by Grant Witness, but curated for us by Tidy Tuesday.

[[NEED TO ADD MORE CONTEXT]]

You can find the data dictionary [here](https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-05-06/readme.md#data-dictionary)

```{r}
nsf_terminations <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-06/nsf_terminations.csv')

```

Let's start with a bit of data cleaning

1. The `org_city` is in all capital letters -- aggro! Edit the `org_city` variable so that cities are in title case instead. (e.g. "Santa Barbara" instead of "SANTA BARBARA")

2. Create a variable `title_length` that gives the number of characters in the grant title. 

3. Create `year_start` and `year_end` variables, which give the *year* that the grant began and the year of the *expected* end date from the original grant. 

4. Create a variable `diversity` that gives the number of times the strings "diversity" or "DEI" appear in the grant abstract. Make sure that this is not case sensitive!

Follow the clues below to find the XX grant.

5. The grant began in 2022 or later. (875 grants remaining)

6. The grant related to indigenous communities, so the word "indigenous" appears in the abstract. Again, pay attention to case! (93 grants remaining)

7. The title was not too long, with less than 100 characters. (Hint: use the `title_length` variable you already created!). (40 grants remaining)

8. The word diversity appears at most one time in the abstract. (35 grants remaining)

9. The start date of a grant was a Wednesday. (2 grants remaining).

10 LAST CLUE to get to FIRE grant

:::callout-note
## Canvas submission

Submit the

:::
