---
title: "Practice Activity 2: Working with R Objects"
subtitle: "Practice with Indexing"
format: html
execute:
  echo: true
  error: true

author: INSTRUCTIONS
eval: false
---
```{r}
#| include: false
#| label: packages

library(tidyverse)
```

Today you will be creating and manipulating vectors, lists, and data frames to
uncover a top secret message.

## Part Zero: Accessing the Practice Activity

Download the template Practice Activity R script here: <a href="../student-version/pa-2-indexing.R" download>`pa-2-indexing.R`</a>

::: callout-important
Be sure to save the file inside your Week 2 folder of your STAT 1810 folder!
:::

## Part One: Debugging

For each question below, you have been given faulty R code. Each code chunk 
will either cause an error or do the desired task incorrectly. Just because
there is no error message doesn't mean the code accomplished what it was 
intended to do! Use the **console** (lower left pane) to check each step and
make sure you have accomplished what you wanted to accomplish.

**Your task is to find the mistake, and correct it so the code completes the intended action.**

1.  Create vectors containing the upper case letters, lower case letters, and
some punctuation marks.

2.  Make **one** long **vector** containing all the symbols.

3.  Turn the `my_symbols` vector into a data frame, with one column named "Symbol".

4.  Find the total number of symbols we have in our data frame.

5.  Create a new variable in your dataframe that assigns a number to each symbol.

## Part Two: Decoding the Secret Message

This chunk will load up the encoded secret message as a vector:

```{r}
#| label: load-secret-data

## we should put this dataset somewhere accessible for everyone (shared ownership)
top_secret <- read_csv("https://www.dropbox.com/s/k72h1zewk4gtqep/PA_Secret_Code?dl=1", 
                       col_names = FALSE)$X1
```

By altering this top secret set of numbers, you will be able to create a
message. Write your own code to complete the steps, **in the order given** below.

For each step, you need to **update the vector after performing an operation.**
Meaning, you need to *overwrite* the existing object with its updated
counterpart (e.g., `x <- x + 12`). 

6.  Add 14 to every number.

7.  Multiply every number by 18, then subtract 257.

8.  Use the `exp()` function to exponentiate every number.

9.  Square every number.

**Checkpoint:** Headquarters has informed you that at this stage of decoding,
there should be 352 numbers in the secret message that are below 17. Write the
code to verify that this is true for your `top_secret` object!

*Hint:* This is what is called a "relational" comparison, where you compare an
object to a number and R will give you a vector of `TRUE`s and `FALSE`s based on
whether the comparison is / is not met. You can then use these `TRUE`s and 
`FALSE`s as numbers, since `TRUE` = `1` and `FALSE` = `0`. 

Next, carry out the following steps:

10. Turn your vector of numbers into a matrix with 5 columns.

11. Separately from your top secret numbers, use the `seq()` function to create
a vector of all the even numbers between 1 and 382. Name this vector `evens`. 
That is, `evens` should contain the numbers 2, 4, 6, 8 ..., 382.

12. Subtract the `evens` vector from the first column of your secret message
matrix. *Remember to save the updated column values in place of the old values!*

13. Subtract 100 from all numbers 18-24th rows of the 3rd column. *Remember to save the updated row values in place of the old values!*

14. Multiply all numbers in the 4th and 5th column by 2. *Remember to save the updated column values in place of the old values!*

15. Turn your matrix back into a vector.

**Checkpoint:** Headquarters has informed you that at this stage of decoding,
all numbers in indices 500 and beyond are below 100. Write the code to verify
that this is true for your `top_secret` object!

*Hint:* Use a relational comparison similar to what you used in the last
checkpoint, but here you will need to subset values from your vector!

16. Take the square root of all numbers in indices 38 to 465. *Remember to save the updated values in place of the old values!*

17. Use the `round()` function to round all numbers to the nearest whole number.

18. Replace all instances of the number 39 with 20. *This requires another relational comparison, but this time it is equality!*

**Checkpoint:** Headquarters has informed you that your final message should
have 344 even numbers. *Checking for divisibility uses the modulus operator (`%%`). Remember numbers that are divisible will have a remainder of 0!*

## Part 3: The secret message!

Use your final vector of numbers as indices for `my_symbols` to discover the
final message, by running the following code:

```{r}
#| label: find-hidden-message
#| error: true

stringr::str_c(my_symbols$Symbol[top_secret], 
               collapse = "")
```

:::callout-tip
# Puzzle Answer

Google the first line of this message to see what poem it is.
:::
