---
title: "Practice Activity 5: Visualize the Tortoises"
format: 
  html:
    code-tools: true
    toc: true      
execute: 
  error: true
  echo: true
  message: false
  warning: false

eval: false
---
# Setup

First, we declare our package dependencies.

```{r}
#| warning: false
#| message: false
#| label: setup

library(readr)
library(ggplot2)
library(dplyr)
```

# Golem Grad Tortoise Data

First, we load in the data. You will have to `install.packages("tidytuesdayR")`.

```{r}
#| label: load-data
tuesdata <- tidytuesdayR::tt_load(2026, week = 9)
tortoise_data <- tuesdata$tortoise_body_condition_cleaned
```

```{r}
#| include: false
# emily exploring...
tortoise_data |>
  ggplot(aes(x = body_mass_grams, y = straight_carapace_length_mm)) +
  geom_point(aes(color = sex), alpha = 0.2)

tortoise_data |>
  ggplot(aes(x = body_condition_index, y = locality)) +
  geom_boxplot(aes(color = sex), alpha = 0.2)

tortoise_data |>
  ggplot(aes(x = as.factor(year_recode), 
             y = body_condition_index)
         ) +
  geom_boxplot()
```

Take a look at the variables in your downloaded data by running the following code. This code with the `glimpse` function reports the data type for each column in the dataset.

```{r}
#| label: preview-data
glimpse(tortoise_data)
```
You should also take a peak at the dataset by double-clicking on it in the environment. 

1. What does one row in the data represent? 

2. What are the some of the other variables and their types? Use `unique()` and `summary()` in the console to explore the data further.

# Select and create an appropriate graphic to answer the following questions:

## "Does it seem easier to capture/recapture male or female tortoises?"

::: tip
How many captures are there for each sex? What variables do you need to answer this question?
:::

3. Use `ggplot2` to draw a barchart of the count of captures for each tortoise sex.

```{r}
tortoise_data |>
  ggplot(aes(x = sex)) +
  geom_bar()
```

## "What are the differences between male and female tortoises across the three locations (Beach Mainland Beach, Plateau Mainland, and Konjsko Island) in terms of body mass or carapace length?"

4. Use `ggplot2` to create a scatterplot of the relationship between the body mass (`body_mass_grams`) and carapace length (`straight_carapace_length_mm`).

::: tip
That's a lot of data! Use `alpha = 0.1` in your scatterplot to create a transparency effect to help with overplotting.
:::

```{r}
tortoise_data |>
  ggplot(aes(x = body_mass_grams, y = straight_carapace_length_mm)) +
  geom_point(alpha = 0.1)
```

5. Building off the previous graph, add an aesthetic to differentiate the location (`locality`) on your scatterplot by color.

```{r}
tortoise_data |>
  ggplot(aes(x = body_mass_grams, 
             y = straight_carapace_length_mm,
             color = sex)
         ) +
  geom_point(alpha = 0.2)
```

5. Building off the same graph, incorporate location (`locality`). There may be more than one method to address this, however, one method will more easily address the questions below.

```{r}
tortoise_data |>
  ggplot(aes(x = body_mass_grams, 
             y = straight_carapace_length_mm,
             color = sex)
         ) +
  geom_point(alpha = 0.2) +
  facet_wrap(~ locality)
```

:::callout-tip
# Canvas Submission
xxx
:::

::: callout-note
Note that this dataset comes from [TidyTuesday 2026 week 9](https://github.com/rfordatascience/tidytuesday/blob/main/data/2026/2026-03-03/readme.md). 
:::
