---
title: "PA 5: Palm Plots"
subtitle: "Data Visualization" 
format: 
  html:
    code-tools: true
    toc: true      
execute: 
  error: true
  echo: true
  message: false
  warning: false

eval: false
---
# Getting Set-up

First, we declare our package dependencies.

```{r}
#| warning: false
#| message: false
#| label: setup

library(readr)
library(ggplot2)
library(dplyr)
```

# Meet the Data

First, we load in the data. You will have to `install.packages("tidytuesdayR")`.

```{r}
#| label: load-data
tuesdata  <- tidytuesdayR::tt_load('2025-03-18')
palmtrees <- tuesdata$palmtrees
```

Take a look at the variables in the `palmtrees` data by running the following code. This code with the `glimpse` function reports the data type for each column in the dataset.

```{r}
#| label: preview-data
tibble::glimpse(palmtrees)
```

You should also take a peak at the dataset by double-clicking on it in the environment.

1.  What does one row in the data represent?

2.  What are the some of the other variables and their types? Use `unique()` and `summary()` in the console to explore the data further.

### Part 1: Fruit Shapes

*Which fruit shapes occur most often?*

3.  Create a bar chart showing the number of palm species with each fruit shape. Add reader-friendly labels

4.  Which two fruit shapes appears most often?

### Part 2: Palm Size

*How are maximum stem height and maximum stem diameter related?*

5.  Create a scatterplot showing the relationship between maximum stem height and maximum stem diameter. Add transparency with `alpha`. Add reader-friendly labels.

Notice that many palm species are compressed into the lower-left corner of the graph because a few species are much larger than the rest. In the Grammar of Graphics, scales control how data values are mapped onto the axes.

6.  Add the following layers to display both axes on a base-10 logarithmic scale.

```         
... +
scale_x_continuous(trans = "log10") +
scale_y_continuous(trans = "log10")
```

7.  What differences do you notice? Why does the log scale make it easier to see the smaller palm species?

### Part 3: Add Subfamily

*Does the relationship between palm height and diameter look different across subfamilies?*

Copy your scatterplot code (with log transformed scales) from above and edit to...

9.  Add color.

10. Try faceting instead.

11. Which version do you prefer to read - color or facets? Why? Does your answer differ based on what you are interested in understanding about the data?

::: callout-tip
# Canvas Submission

1.  Which **two** fruit shapes occur most often?
2.  Which palm subfamily appears to include the tallest species while also having relatively small stem diameters?
3.  Why don't palm species with a stem height or diameter of **0** appear 'nicely' on the log-transformed scatterplot? \*(Hint: What is (\log\_{10}(0))?)\*
4.  Which palm subfamily appears to contain only a single species?
:::
