---
title: "PA 8: Twinning"
subtitle: "Data Joins"
format: 
  html:
    code-tools: true
    toc: true      
execute: 
  error: true
  echo: true
  message: false
  warning: false

eval: false
---
```{r setup}
# load packages
library(readxl) 
library(tidyverse)
```

The data come to us from [TidyTuesday](https://github.com/rfordatascience/tidytuesday/blob/main/data/2026/2026-05-12/readme.md#data-dictionary) -- originally compiled by 
[Ahmad Barclay](https://github.com/bothness/twin-cities) and further curated by [Nicola Rennie](https://github.com/nrennie) for TidyTuesday.

We also created a dataset `dists` with the following columns:

- `source`: same as in `links`
- `target`: same as in `links`
- `dist`: distance between paired cities in km

```{r}
cities <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-05-12/cities.csv')
links <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-05-12/links.csv')

# NEED TO CHANGE THIS TO A LINK TO OUR GITHUB
dists <- read_csv("../../data/08-dists.csv")
```

1. Check out the data. What is one observation in the `cities` dataset and what kind of information do we get from this data? What is one observation in the `links` dataset and what kind of information do we get from this data?

The `links` data contains all of the pairs of cities, but we can't learn anything interesting just from the city IDs! We want to add the city information on.

2. What are the key(s) between the `city` and `links` data?

As the data dictionary states, the order of "source" and "target" city is arbitrary. 

3. Created a new dataframe called  `linked_cities` that includes all of the information in the `cities` data for both cities in each pair of twinned cities.

Note: This means you will have *two* name, latitude, longitude, etc. columns for each row. Think about how you might want to name columns to keep things from being confusing!

4. Are any of the twinned cities missing city information like their name?

```{r}
linked_cities |> 
  filter(is.na(source_name) | is.na(name))
```

5. Q4 implies that we could have used which of the following joins (or all of them)? inner, left, right, and/or full?

6.  Join the distances between cities on to your `linked_cities` data. Pay close attention: what identifies one observation in the `dists` data? Your `linked_cities` data should still have 10,596 observations.

7. Which two twin cities are the *farthest* apart?

8. Which two twin cities are the *closest* together? Is this a mistake in the data!? Look up the two cities on a map.

9. Which country has the most twinned cities within the same country?

10. CHALLENGE: Which continent has the most twinned cities with North America?

```{r}
linked_cities |> 
  filter(continent_1 == "North America"|
           continent_2 == "North America") |> 
  group_by(continent_1, continent_2) |> 
  count() |> 
  pivot_longer(starts_with("continent"),
               names_to = "order",
               values_to = "continent") |> 
  filter(continent != "North America") |> 
  summarize(n = sum(n), .by = continent) |> 
  arrange(desc(n))
```
