---
title: "PA 9: Cancel Culture"
subtitle: "Strings and Date Variables"
filters:
  - versions
versions:
  - student:
      out-dir: ".."
      yaml:
        author: "Your Name"
        embed-resources: true
  - instructions:
      out-dir: ".."
      yaml:
        eval: false
  - solution:
      out-dir: ".."
      out-file: "pa9-solution.qmd"
      yaml:
        author: "SOLUTIONS"
        embed-resources: true
format: 
  html:
    code-tools: true
    toc: true      
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

```{r setup}
# load packages
library(tidyverse)
```

The data are compiled by Grant Witness, but curated for us by Tidy Tuesday.

[[NEED TO ADD MORE CONTEXT]]

You can find the data dictionary [here](https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-05-06/readme.md#data-dictionary)

```{r}
nsf_terminations <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-06/nsf_terminations.csv')

```


Let's start with a bit of data cleaning

1. The `org_city` is in all capital letters -- aggro! Edit the `org_city` variable so that cities are in title case instead. (e.g. "Santa Barbara" instead of "SANTA BARBARA")


```{r}
#| version: solution

nsf_terminations <- nsf_terminations |> 
  mutate(org_city = str_to_title(org_city))
```


2. Create a variable `title_length` that gives the number of characters in the grant title. 

```{r}
#| version: solution

nsf_terminations <- nsf_terminations |> 
  mutate(title_length = str_length(project_title))
```


3. Create `year_start` and `year_end` variables, which give the *year* that the grant began and the year of the *expected* end date from the original grant. 

```{r}
#| version: solution

nsf_terminations <- nsf_terminations |> 
  mutate(year_start = year(nsf_startdate),
         year_end = year(nsf_expected_end_date))
```

4. Create a variable `diversity` that gives the number of times the strings "diversity" or "DEI" appear in the grant abstract. Make sure that this is not case sensitive!

```{r}
#| version: solution

nsf_terminations <- nsf_terminations |> 
  mutate(diversity = str_count(str_to_lower(abstract),
                               "diversity|dei"))
```

Follow the clues below to find the XX grant.


5. The grant began in 2022 or later. (875 grants remaining)

```{r}
#| version: solution

clues <- nsf_terminations |> 
  filter(year_start >= 2022)
```

6. The grant related to indigenous communities, so the word "indigenous" appears in the abstract. Again, pay attention to case! (93 grants remaining)

```{r}
#| version: solution

clues <- clues |> 
  filter(str_detect(str_to_lower(abstract), "indigenous"))
```

7. The title was not too long, with less than 100 characters. (Hint: use the `title_length` variable you already created!). (40 grants remaining)

```{r}
#| version: solution

clues <- clues |> 
  filter(title_length < 100)
```

8. The word diversity appears at most one time in the abstract. (35 grants remaining)

```{r}
#| version: solution

clues <- clues |> 
  filter(diversity <= 1)
```


9. The start date of a grant was a Wednesday. (2 grants remaining).

```{r}
#| version: solution

clues <- clues |> 
  filter(wday(nsf_startdate) == 4)
```

10 LAST CLUE to get to FIRE grant



:::callout-note
## Canvas submission

Submit the

:::
