---
title: "PA 9: Cancel Culture"
subtitle: "Strings and Date Variables"
format: 
  html:
    code-tools: true
    toc: true      
execute: 
  error: true
  echo: true
  message: false
  warning: false

eval: false
---
```{r setup}
# load packages
library(tidyverse)
library(knitr)
```

[[NEED MORE PREAMBLE]]

The data are compiled by Grant Witness.

You can find the data dictionary [here](https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-05-06/readme.md#data-dictionary)

```{r}
nsf_terminations <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-06/nsf_terminations.csv')

```

## Data Context

First, as always, let's spend some time getting acquainted with the data. 

1. How does this data relate to you and your life? What are your impressions of the social and political context of this data? What questions might you want to answer using this data?

2. Looking at the data and the data dictionary together, do you have any questions about what any of the columns represent? 

3. What does one row of the data represent? What is included in the sample -- are any possible observations in a full population left out?

4. Which variables in the data represent dates? Are they stored as dates in the `nsf_terminations` data, or do we need to convert them to dates? What is the format? If you need to convert them, do that.

## Data Cleaning

Next, let's clean up the data a little bit and adding some variables that we will use later.

Implement the following in one dplyr pipeline. But go one step at a time to catch any bugs as you go!

5. The city name is annoyingly mostly in all caps but sometimes in title case -- edit the `org_city` variable so that it is all in title case.

6. Create two new variables that give the year that the grant started and the intended end year of the grant. Call these variables `start_year` and `expected_end_year`.

7. Create a variable called `diversity` that gives the *number of times* that the word "diversity" or "equity" appears in the grant abstract. Be careful - make sure this does not depend on case!

## Find that Grant

Later this week, we will explore general trends in the cancelled grants more. However, each grant tells a whole story and represents many people and resources that were impacted by the termination.

Follow the clues below to narrow our investigation down to one grant.

<!--Add more context around why we are looking at one grant?? There is really no reason to search for this grant in particular... I just thought it was interesting and worked with some relevant skills to narrow it down -->

8. The grant number starts with "23". Filter to these grants using one function from `stringr`. (Hint: do you need to change the variable type to do this?). 

9. The grant was intended to span four years or more. (For example, a grant that started in the year 2021 and was expected to end in 2025 spanned 5 years.)

10. The grant we are looking for involves indigenous communities, so the word "indigenous" appears in the abstract. Be careful to make this search not case-sensitive!

11. The title of the grant isn't tooooo long - it is  not more than 120 characters.

12. The start date of the grant was a Friday.

13. Finally, the word "diversity" does appear in the abstract, but even so this grant was not on Ted Cruz's infamous list.

14. Take a look at this grant -- is our data out of date and it was actually reinstated?
(Hint: you can try searching the title or NSF grant number in a search engine, or you can look for it on the updated [Grant Witness page](https://grant-witness.us/nsf-data.html).)

:::callout-note
## Canvas Submission

Submit the amount of money that was still obligated to the grantees when the grant was terminated and the organizations city and state as they appear in your cleaned data. 

:::
