---
title: "Practice Activity 2: Working with R Objects"
subtitle: "Practice with Indexing"
author: "Solutions"
format: html
execute:
  echo: true
---

## Part One: Debugging

For each question below, you have been given faulty R code. Each code chunk 
will either cause an error or do the desired task incorrectly. Just because
there is no error message doesn't mean the code accomplished what it was 
intended to do! Use the **console** (lower left pane) to check each step and
make sure you have accomplished what you wanted to accomplish.

**Your task is to find the mistake, and correct it so the code completes the intended action.**

1.  Create vectors containing the upper case letters, lower case letters, and some punctuation marks.

```{r}
#| label: question-1

lower_case <- c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j", "k", "l", "m",
                "n", "o", "p", "q", "r", "s", "t", "u", "v", "w", "x", "y", "z")

## missing comma after "H"
upper_case <- c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K", "L", "M",
                "N", "O", "P", "Q", "R", "S", "T", "U", "V", "W", "X", "Y", "Z")

## needs a '' around a " character
punctuation <- c(".", ",", "!", "?", "'", '"', "(", ")", " ", "-", ";", ":")

```

2.  Make **one** long **vector** containing all the symbols.

```{r}
#| label: question-2

## The c() function makes a vector, not cbind()
my_symbols <- c(lower_case, upper_case, punctuation)

```

3.  Turn the `my_symbols` vector into a data frame, with one column named "Symbol".

```{r}
#| label: question-3

## The correct function name is data.frame (not dataframe)
my_symbols <- data.frame(Symbol = my_symbols)
```

4.  Find the total number of symbols we have in our data frame.

```{r}
#| label: question-4

## The correct function name is nrow() (not nrows)
len <- nrow(my_symbols)
```

5.  Create a new variable in your dataframe that assigns a number to each symbol.

```{r}
#| label: question-5

## To make a sequence of numbers you use a : between them (not a %)  
my_symbols$Num <- 1:len
```

## Part Two: Decoding the Secret Message

This chunk will load up the encoded secret message as a vector:

```{r}
#| label: load-secret-data

library(tidyverse)

top_secret <- read_csv("https://www.dropbox.com/s/k72h1zewk4gtqep/PA_Secret_Code?dl=1", 
                       col_names = FALSE)$X1
```

By altering this top secret set of numbers, you will be able to create a
message. Write your own code to complete the steps, **in the order given** below.

For each step, you need to **update the vector after performing an operation.**
Meaning, you need to *overwrite* the existing object with its updated
counterpart (e.g., `top_secret[2] <- top_secret[2] + 12`). 

6.  Add 14 to every number.

```{r}
#| label: question-6

top_secret <- top_secret + 14
```

7.  Multiply every number by 18, then subtract 257.

```{r}
#| label: question-7

top_secret <- (top_secret * 18) - 257
```

8.  Use the `exp()` function to exponentiate every number.

```{r}
#| label: question-8

top_secret <- exp(top_secret)
```

9.  Square every number.

```{r}
#| label: question-9

top_secret <- top_secret^2
```

**Checkpoint:** Headquarters has informed you that at this stage of decoding,
there should be 352 numbers in the secret message that are below 17. Write the
code to verify that this is true for your `top_secret` object!

```{r}
#| label: checkpoint-1

sum(top_secret < 17)

```

Next, carry out the following steps:

10. Turn your vector of numbers into a matrix with 5 columns.

```{r}
#| label: question-10

top_secret <- matrix(top_secret, ncol = 5)
```

11. Separately from your top secret numbers, create a vector of all the even
numbers between 1 and 382. Name it `evens`. That is, `evens` should contain 
the numbers 2, 4, 6, 8 ..., 382.

```{r}
#| label: question-11

evens <- seq(from = 2, to = 382, by = 2)
```

12. Subtract the `evens` vector from the first column of your secret message
matrix. 

```{r}
#| label: question-12

top_secret[, 1] <- top_secret[, 1] - evens

```

13. Subtract 100 from all numbers 18-24th rows of the 3rd column. 

```{r}
#| label: question-13

top_secret[18:24, 3] <- top_secret[18:24, 3] - 100

```

14. Multiply all numbers in the 4th and 5th column by 2. 

```{r}
#| label: question-14

top_secret[, 4:5] <- top_secret[, 4:5] * 2

```

15. Turn your matrix back into a vector.

```{r}
#| label: question-15

top_secret <- c(top_secret)
```

**Checkpoint:** Headquarters has informed you that at this stage of decoding,
all numbers in indices 500 and beyond are below 100. Write the code to verify
that this is true for your `top_secret` object!

```{r}
#| label: checkpoint-2

sum(top_secret[500:955] < 100)

length(500:955)
```

16. Take the square root of all numbers in indices 38 to 465. 

```{r}
#| label: question-16

top_secret[38:465] <- sqrt(top_secret[38:465])
```

17. Use the `round()` function to round all numbers to the nearest whole number.

```{r}
#| label: question-17

top_secret <- round(top_secret)
```

18. Replace all instances of the number 39 with 20. 

```{r}
#| label: question-18

top_secret[top_secret == 39] <- 20
```

**Checkpoint:** Headquarters has informed you that your final message should
have 344 even numbers. *Checking for divisibility uses the modulus operator (`%%`). Remember numbers that are divisible will have a remainder of 0!*

```{r}
#| label: checkpoint-3

sum(top_secret %% 2 == 0)
```

## Part 3: The secret message!

Use your final vector of numbers as indices for `my_symbols` to discover the
final message, by running the following code:

```{r}
#| label: find-hidden-message

stringr::str_c(my_symbols$Symbol[top_secret], 
               collapse = "")
```

:::callout-tip
# Puzzle Answer

Jabberwocky by Lewis Carroll
:::
