---
title: "Practice Activity 3: Control Flow"
subtitle: "Practice with Loops & Conditionals"
format: html
execute:
  echo: true
  error: true

author: SOLUTIONS
eval: true
---
```{r}
#| include: false
#| label: packages

library(tidyverse)
```

The ba`R`istas at Tidy `R`oasters have received a huge drink order and are
preparing to fill it. Your job is to help summarize, and fix, the orders.

## Part One: Getting the Drink Orders

1.  Let's get the drink order data into `R`. Since we haven't covered reading
in data yet, we'll make it easy! Follow these steps.

    1.  In RStudio, go to the *Files* pane and navigate to the saved data.

    2.  Click on the file name and select *Import Dataset...*.
    
    3.  Keep all of the settings and click *Import*.
    
You should now see the `coffee_orders` dataset in your Environment tab (upper
right corner). 

```{r}
#| echo: false
#| label: read-in-data

coffee_orders <- readxl::read_xlsx(
  here::here("practice-activities", 
             "student",
             "coffee_orders.xlsx")
  )
```

```{r}
#| echo: true
#| label: preview-data

coffee_orders
```

## Part Two: Getting Information About the Drink Orders

2.  How many drinks are in the order? Save this number in an object named 
`num_orders` and print out the value.

```{r}
#| label: q2
#| echo: true

num_orders <- nrow(coffee_orders)
num_orders
```

3. Think back to last week. What operator allows you to grab a single column
from a dataframe? Using that operator, write a line of code that produces a 
vector of the drink `Base` for all drinks. 

```{r}
#| label: q3
#| echo: true

coffee_orders$Base
```

4. Now write a line of code that produces the drink `Base` for the fourth drink
order.

```{r}
#| label: q4
#| echo: true

coffee_orders$Base[4]
```

## Part Three: Finding Specific Drink Orders

5. Use a `while()` loop to find the first person to order a drink with a chai
`Base`. Report this person's full drink order.

```{r}
#| label: q5
#| echo: true

base <- ""
counter <- 0
while (base != "chai") {
  counter <- counter + 1
  base <- coffee_orders$Base[counter]
}
coffee_orders[counter,]
```

6. Use a `for()` loop to count the total number of espresso drinks. 

```{r}
#| label: q6
#| echo: true

num_espresso <- 0

for(i in 1:num_orders){
  if (coffee_orders$Base[i] == "espresso") {
    num_espresso <- num_espresso + 1
  }
}
num_espresso
```

## Part Four: Making Drink Modifications

7. Oh no! Tidy `R`oasters has run out of soy milk. The ba`R`istas decide to use
oat milk instead. In the `coffee_orders` dataframe, change all the soy milk 
requests to oat milk. Use the `if_else()` function from `dplyr` to do this.

```{r}
#| label: q7
#| echo: true

coffee_orders$Milk <- if_else(condition = coffee_orders$Milk == "soy",
                              true = "oat",
                              false = coffee_orders$Milk)
```

8. The ba`R`istas need to know how much ice to have on hand. For a medium iced
drink, they need 4oz of ice. For a large iced drink, they need 6oz of ice. Use 
a `for()` loop and conditionals to calculate the total ice amount the ba`R`istas
need.

```{r}
#| label: q8

ice <- 0
for (i in 1:num_orders) {
  if (coffee_orders$Temperature[i] == "iced") {
    if (coffee_orders$Size[i] == "medium") {
      ice <- ice + 4
    }
    else {
      ice <- ice + 6
    }
  }
}
ice
```
