---
title: "Week 2 Code Along"
filters:
  - versions
versions:
  - instructions:
      out-dir: ".."
      yaml:
        eval: false
        echo: true
        author: "Instructions"
  - solution:
      out-dir: ".."
      out-file: 02-code-along-solution.qmd
      yaml:
        author: "SOLUTIONS"
        eval: 
        error: true
format: 
  html:
    toc: true      
---

Let's consider the top five grossing movies in the first half of 2026. We'll create a dataset that contains...

(a) the movie name,
(b) the total amount grossed (in millions of dollars), and
(c) whether or not the movie is part of a larger franchise.

## Creating a Dataframe

1. Open a new R script.

2. Copy this code to create a **dataframe** called `movie_info`. The dataframe
should have 3 columns: `name`, `gross`, and `franchise`, and five total movies 
(rows).

```{r}
movie_info <- data.frame(name = c("The Super Mario Galaxy Movie",
                                  "Michael",
                                  "The Devil Wears Prada 2",
                                  "Project Hail Mary",
                                  "Toy Story 5"),
                         gross = c(1009, 
                                   978, 
                                   686, 
                                   682, 
                                   610),
                         franchise = c(TRUE,
                                       FALSE,
                                       TRUE,
                                       FALSE,
                                       TRUE)
                         )
```

Let's **look at** the dataframe we just created!

3. Go to your *Console*, type `movie_info`, and hit enter.

4. Go to the *Environment* pane, find `movie_info`, and click on it.

5. What code appears in the *Console*? This is the code you could have run 
to open the data viewer! 

```{r}
#| version: solution

View(movie_info)
```


## Dataframe Indexing

Often, we want to work with a **single column** (variable) or a **single row**
(observation) in a dataframe. There are many ways to do this!

6. Use `[ , ]` notation to pull out the vector of gross revenue.

```{r}
#| version: solution

movie_info[,2]
```

7. Use `$` notation to pull out the vector of gross revenue.

```{r}
#| version: solution

movie_info$gross
```

8. Use `[ , ]` notation to pull out the data on *Project Hail Mary*.

```{r}
#| version: solution

movie_info[4,]
```

9. What proportion of the top five grossing movies are part of a larger
franchise?

```{r}
#| version: solution

sum(movie_info$franchise) / nrow(movie_info)
```

## Data Types

A benefit of dataframes, and why they are particularly useful in statistics, is
that they can contain columns of different data types. Let's use the `typeof()`
and `class()` functions to obtain information about the structure and data types
of the `movie_info` object. 

10. What does the `typeof()` function return when you input `movie_info`?
Why do you get this output?

```{r}
#| version: solution

typeof(movie_info)
```

::: {.version-solution}
The `typeof()` function tells us how `movie_info` is stored. Although we created
`movie_info` as a dataframe, behind the scenes a dataframe behaves the same as
a list. 
:::

10. What does the `class()` function return when you input `movie_info`?
Why do you get this output?

```{r}
#| version: solution

typeof(movie_info)
```

::: {.version-solution}
The `class()` function also tells us how `movie_info` is stored. However, the
`class()` function is more specific. It tells us that `movie_info` is a 
dataframe, a specific type of list. 
:::

12. Use `typeof()` to check the data type of each column within `movie_info`. 
You'll have to pull out each column to check!

```{r}
#| version: solution

typeof(movie_info$name)
typeof(movie_info$gross)
typeof(movie_info$franchise)

```

13. Use `class()` to check the data type of each column within `movie_info`. 
How do these outputs differ from the results you obtained from `typeof()`?

```{r}
#| version: solution

class(movie_info$name)
class(movie_info$gross)
class(movie_info$franchise)

```

::: {.version-solution}
The output is the same except for `class(movie_info$gross)`. For the `typeof()`
function the output was `"double"` and for the `class()` function the output was 
`"numeric"`. What we are seeing is the opposite of what we saw
before---`typeof()` is giving us a more specific description (double) and
`class()` is giving us a more general description (numeric).
:::
