---
title: "Week 3 Code Along"
filters:
  - versions
versions:
  - instructions:
      out-dir: ".."
      yaml:
        eval: false
        echo: true
        author: "Instructions"
  - solution:
      out-dir: ".."
      out-file: 03-code-along-solution.qmd
      yaml:
        author: "SOLUTIONS"
        eval: 
        error: true
format: 
  html:
    toc: true      
---


You're a new employee at a non-profit committed to sustainability. The company
has recently pledged to plant trees to offset their paper use.

Since the company places several paper orders throughout the year, you have been
asked to **write a function** to calculate the number of trees to plant based on
a given order. There are three types of paper used by the company, each of which
requires a different number of trees. One carton (5000 sheets) of...

- office paper uses 0.6 trees.
- newsprint (for the newsletter) uses 0.3 trees.
- glossy paper (for the catalog) uses 0.4 trees.

In order to offset waste, the company wants to plant 125% of the trees that
directly go into the paper they buy.

1. Write a function named `trees_to_plant()` that takes in two arguments

- the number of cartons (`cartons`)
- the material (`material`) that was ordered

The function should output the number of trees the company needs to plant.

```{r}
#| version: solution

trees_to_plant <- function(cartons, material){
  if(material == "office"){
    num_trees <- 0.6 * cartons
  }
  else if(material == "newsprint"){
    num_trees <- 0.3 * cartons
  }
  else {
    num_trees <- 0.4 * cartons
  }
  num_trees <- num_trees * 1.25
  return(num_trees)
}
```

2. Run the function. Does anything happen?

::: {.version-solution}
Nothing happens when you run the function!
:::

3. Take a look at the *Environment* pane. Do you notice anything there?

::: {.version-solution}
Although nothing is output when you run the function, the `trees_to_plant()`
function appears in the Environment. `trees_to_plant()` is listed under the 
"Functions" objects that have been created and the Environment shows that the 
function takes two inputs `(cartons, material)`.
:::

Nothing really happens when you **run** the code that defines a function. The
action happens when you **call** that function!

4. The non-profit places an order for 5 cartons of office paper. Run your
function to find the number of trees to plant.

```{r}
#| version: solution

trees_to_plant(cartons = 5, material = "office")
```

5. How many trees for 2 cartons of glossy paper?

```{r}
#| version: solution

trees_to_plant(cartons = 2, material = "gloss")
```

The function is doing lots of work behind the scenes. **But, what happens inside
the function is not accessible outside the function!**

6. Inside our function, the object `num_trees` was created. Can we see 
that object now? Try to print out `num_trees` in the *Console*.

```{r}
#| version: solution
#| error: true

num_trees
```

7. Define `num_trees <- 7`. Then run `trees_to_plant()` for 1 carton of 
newsprint. What is the value of `num_trees` now?

```{r}
#| version: solution
num_trees <- 7

trees_to_plant(carton = 1, material = "newsprint")
```

