---
title: "Code-Along 7: More Data Wrangling"
filters:
  - versions
versions:
  - student:
      out-dir: ".."
      yaml:
        author: "Your Name"
        embed-resources: true
  - solution:
      out-dir: ".."
      out-file: "07-code-along-solution.qmd"
      yaml:
        author: "SOLUTIONS"
        embed-resources: true
format: 
  html:
    code-tools: true
    toc: true      
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

```{r}
#| label: setup
#| echo: false

library(tidyverse)
library(liver)
library(knitr)
data(cereal)
```


## Revisit Plot

Remember the plot we started working on this week? Let's actually create it! Follow the steps below to build up the code.


```{r}
#| version: student

cereal |> 
  select(-weight, -cups, -rating) 

```

```{r}
#| version: solution

cereal |> 
  select(-weight, -cups, -rating) |> 
   pivot_longer(cols = calories:vitamins,
               names_to = "nutrient",
               values_to = "amount") |> 
  group_by(shelf, nutrient) |> 
  summarise(mean_amount = mean(amount)) |> 
  ggplot(aes(x = shelf, 
             y = mean_amount, 
             color = nutrient)) +
  geom_point() +
  geom_line() +
  labs(x = "Shelf", 
       y = "", 
       subtitle = "Mean Amount",
       color = "Nutrient")

```


1. The first step is completed for you -- removing some columns we don't need.

2. Reshape the data **longer**. You should have this code written down in handout 7!

3. Check out the long data. Does it look how you expected? What does one row represent? Did you draw it correctly in handout 7?

:::{.version-solution}
*One row is one cereal and nutrient combination! There are 9 nutrients, so there are 9 rows per cereal. Since the name, manufacturer, type, and shelf are unique for a cereal, these values are repeated for each of the 9 rows per cereal.*

:::

4. The plot shows the mean nutrient amount for each nutrient and shelf. Calculate these values using your dplyr skills.

5. Create the plot! 


## More Reshape

Tables can be useful to communicate information from data as well! 

Everyone is obsessed with protein and fiber these days! 

**Goal**: A table of the number of cereals that are low/medium/or high protein and low or high fiber. 

1. Create a variable called `protein_level` labeling each cereal as "low", "medium", or "high" protein, using the following categorizations:

 - low: 0-1 g protein
 - medium: 2-4 g protein
 - high: more than 4 g protein
 
2. Create a variable called `fiber_level` labeling each cereal as "low" or "high" fiber, using the following categorizations:

 - low: 0-2 g fiber
 - high: more than 2 g fiber
  
3. Find the number of cereals for each combination of protein and fiber level.

4. Is this table easy to read? What might help?

```{r}
#| version: student

cereal |> 
  
```


```{r}
#| version: solution

cereal |> 
  mutate(protein_level = case_when(protein <= 1 ~ "low",
                                   protein <= 4 ~ "medium",
                                   .default = "high"),
         fiber_level = if_else(fiber <= 2 ,
                               "low",
                               "high")) |> 
  count(protein_level, fiber_level) |> 
  pivot_wider(values_from = n,
              names_from = fiber_level)
```
5. Reshape the data so that there is a column for each fiber level.


6. What do you think now? What would help make this table more readable?


7. One way to present rows of dataframes nicely when you render your Quarto is to use the `kable()` function in the knitr package. Format your output nicely with `kable()`.
