---
title: "Code-Along 9: Strings and Dates"
filters:
  - versions
versions:
  - student:
      out-dir: ".."
      yaml:
        author: "Your Name"
        embed-resources: true
  - solution:
      out-dir: ".."
      out-file: "09-code-along-solution.qmd"
      yaml:
        author: "SOLUTIONS"
        embed-resources: true
format: 
  html:
    code-tools: true
    toc: true      
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

```{r}
nsf_terminations <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-06/nsf_terminations.csv')
```

## More Dates - Periods

**Goal:** Calculate the percent of the time of the grant that was taken away by it being terminated.

1. Calculate the intended total length of the grant in **days**. 

2. Calculate the number of days that should have remained of the grant after the termination letter was sent.

3. Use these two variables to calculate the percent of days taken from the grant recipients. 

```{r}
#| version: solution

nsf_terminations <- nsf_terminations |> 
  mutate(intended_length = as.period(nsf_expected_end_date - 
                                       nsf_startdate),
         missing_length = as.period(nsf_expected_end_date -
                                      termination_letter_date),
         lost_perc = missing_length / intended_length)
```

4. What was the average and total amount of lost time by state in terms of years?

```{r}
#| version: solution

nsf_terminations |> 
  group_by(org_state) |> 
  summarize(mean_years = mean(missing_length / years(1)),
            total_years = sum(missing_length / years(1)),
            n = n()) |> 
  arrange(desc(total_years))
```



5. How many grants were complete before they were terminated?

```{r}
#| version: solution

nsf_terminations |> 
  filter(lost_perc <= 0) |> 
  nrow()
```


## Strings - `str_replace()`

6. 

- "clean up" the grant name by replacing words `str_replace_all()`
- `str_extract()`

