---
title: "Code-Along 9: Strings and Dates"
filters:
  - versions
versions:
  - student:
      out-dir: ".."
      yaml:
        author: "Your Name"
        embed-resources: true
  - solution:
      out-dir: ".."
      out-file: "09-code-along-solution.qmd"
      yaml:
        author: "SOLUTIONS"
        embed-resources: true
format: 
  html:
    code-tools: true
    toc: true      
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

```{r}
#| label: setup

library(tidyverse)
library(readxl)
```


```{r}
nsf_terminations <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-06/nsf_terminations.csv')
```

## More Dates - Periods

**Goal:** Calculate the percent of the time of the grant that was taken away by it being terminated.

1. Calculate the intended total length of the grant in **days**. 

2. Calculate the number of days that should have remained of the grant after the termination letter was sent.

3. Use these two variables to calculate the percent of days taken from the grant recipients. 

```{r}
#| version: student

# Q1-3 code
```


```{r}
#| version: solution

nsf_terminations <- nsf_terminations |> 
  mutate(intended_length = as.period(nsf_expected_end_date - 
                                       nsf_startdate),
         missing_length = as.period(nsf_expected_end_date -
                                      termination_letter_date),
         lost_perc = missing_length / intended_length)
```

4. What was the average and total amount of lost time by state in terms of years?

```{r}
#| version: solution

nsf_terminations |> 
  group_by(org_state) |> 
  summarize(mean_years = mean(missing_length / years(1)),
            total_years = sum(missing_length / years(1)),
            n = n()) |> 
  arrange(desc(total_years))
```

```{r}
#| version: student

# Q4 code
```


5. How many grants were complete before they were terminated?

```{r}
#| version: solution

nsf_terminations |> 
  filter(lost_perc <= 0) |> 
  nrow()
```

```{r}
#| version: student

# Q5 code
```

## Strings

One of the variables in our data is `in_cruz_list`, which is a reference to a [report](https://democrats-science.house.gov/imo/media/doc/democratic_staff_report-defending_hidden_figures.pdf) by Ted Cruz, a US senator from Texas, which labeled 3,483 NSF grants as  "questionable projects that promoted diversity, equity, and inclusion (DEI) tenets or pushed onto science neo-Marxist perspectives about enduring class struggle." There is also a [rebuttal report](https://www.govinfo.gov/content/pkg/GOVPUB-Y4_C73_7-PURL-gpo234941/pdf/GOVPUB-Y4_C73_7-PURL-gpo234941.pdf). 

The report lists a large number of key words that the Cruz report used to identify these grants. Let's see if we can find these terms in the cancelled grants. We compiled the terms into an excel sheet to make things easier.

```{r}
search_terms <- read_xlsx("../../data/09-cruz_terms.xlsx")
```


6. In order to figure out if any one of these terms is found in the grant title or abstract, we need to condense the `term` column in `search_terms` into a **single** string, call it, `term_string` where...

a. each term is all lowercase
b. each term is separated by a pipe "|"

```{r}
#| version: solution

term_string <- search_terms |> 
  mutate(term = str_to_lower(term)) |> 
  pull() |> 
  str_flatten(collapse = "|")
```

```{r}
#| version: student

# Q6 code
```

Now `term_string` is a SUPER long string:

```{r}
str_length(term_string)
```

7. Use `term_string` to create an indicator variable called `term_found` 
in `nsf_terminations` that takes the value 1 if one of the Cruz terms is 
found in the title or abstract and 0 otherwise.

```{r}
#| version: student

# Q7 code
```

```{r}
#| version: solution

nsf_terminations <- nsf_terminations |> 
  mutate(term_found = case_when(str_detect(str_to_lower(project_title), term_string) ~ 1,
                                str_detect(str_to_lower(abstract), term_string) ~ 1,
                                .default = 0))
```

8. Use `term_string` to create a variable called `first_term`, that gives
the first of the Cruz terms that appeared in the grant's abstract. 

```{r}
#| version: student

# Q8 code
```

```{r}
#| version: solution

nsf_terminations <- nsf_terminations |> 
  mutate(first_term = str_extract(str_to_lower(abstract), term_string))
```

9. What are the top 10 terms that were found first?

```{r}
#| version: student

# Q9 code
```

```{r}
#| version: solution

nsf_terminations |> 
  count(first_term) |> 
  slice_max(order_by = n, n = 10)
```


10. Help the researchers out by replacing the term "historically" with "previously" in their abstract so that they won't be flagged next time! Check that it worked!

```{r}
#| version: student

# Q10 code
```

```{r}
#| version: solution

nsf_terminations <- nsf_terminations |> 
  mutate(abstract = str_replace_all(abstract, "historically", "previously"),
         abstract = str_replace_all(abstract, "Historically", "Previously"))

nsf_terminations |> 
  filter(str_detect(abstract, "historically|Historically")) |> 
  nrow()
```

