---
title: "Lab 3: Conditionals"
filters:
  - versions
versions:
  - instructions:
      out-dir: ".."
      yaml:
        eval: true
        echo: true
        author: "Instructions"
  - solution:
      out-dir: ".."
      out-file: "lab3-solution.qmd"
      yaml:
        author: "SOLUTIONS"
        eval: true
        error: true
format: 
  html:
    toc: true      
---

It's been hotly contested, but your R instructors have named the (controversial) **best breakfast burrito in SLO**! To discover the winner, decode their message.

We will break this task down into several small steps and write a function for
each.

:::{.version-instructions}
::: {.callout-tip collapse="true"}
## Seeking Help

Part of learning to program is learning from a variety of resources. Thus, I expect you will use resources that you find on the internet. There is, however, an important balance between copying someone else's code and *using their code to learn*. Therefore, if you use external resources, I want to know about it. You can "inform" me of any resources you used by **pasting the link to the resource in a code comment next to where you used that resource**.

Additionally, you are permitted and encouraged to work with your peers as you complete lab assignments, but **you are expected to do your own work**. Copying from each other is cheating, and letting people copy from you is also cheating. Please don't do either of those things.
:::
:::

:::{.version-instructions}
## Part 0: Accessing the Lab

You can download the R script for Lab 3 here: <a href="../student/lab3-functions-control-flow.R" download>`lab3-functions-control-flow.R`</a>
:::

## Part 1: Write and Test Your Functions

### Function 1: `divide_and_round()`

**1. Write code to manipulate the vector of numbers `nums`. Your code should divide each element in the vector by the smallest element and then round the results to the nearest whole number.** *This code should not be a function.*

```{r}
nums <- 3:12
# Add your code below.
```

:::{.version-solution}
```{r}
min_elm <- min(nums, na.rm = TRUE)
divided <- nums/min_elm
round(divided)
```
:::

**2. Turn your code from Q1 into a function called `divide_and_round()`.** *Fill in the skeleton below.*

```{r}
# Complete the function.
divide_and_round <- function(vec){

}
```

:::{.version-solution}
```{r}
divide_and_round <- function(vec){
  min_elm <- min(vec, na.rm = TRUE)
  divided <- vec/min_elm
  rounded <- round(divided)
  return(rounded)
}
```
:::

**3. Test your function by running the code below.** Make sure the output is as expected. You may need to go back to Q2 and make changes to your function!

```{r}
# Don't change this code!
test <- c(5:10, NA)
divide_and_round(test)
```

**4. Why is it important to test the functions that you write? Did the test in Q3 help you improve your function?**

### Function 2: `no_fours_or_fourteens()`

**5. Write code to manipulate the vector of numbers `nums`. Your code should return a vector of `TRUE` and `FALSE`. For each element, return `TRUE` if the number is NOT divisible by 4 or 14, and return `FALSE` otherwise.** 
*This code should not be a function.*

:::{.version-solution}
```{r}
!(nums %% 4 == 0| nums %% 14 == 0)
```
:::

**6. Turn your code from Q5 into a function called `no_fours_or_fourteens()`.**

:::{.version-solution}
```{r}
no_fours_or_fourteens <- function(vec){
  check <- !(vec %% 4 == 0| vec %% 14 == 0)
  return(check)
}
```
:::

**7. Test your function by running the code below.** Make sure the output is as expected. You may need to go back to Q6 and make changes to your function!

```{r}
# Don't change this code!
test <- c(3:14*5, NA)
no_fours_or_fourteens(test)
```

### Function 3: `every_other_num()`

**8. Write a line of code that returns `TRUE` if `value` is odd and `FALSE` is `value` is even. Test your code on `value <- 3` and `value <- 6`.**

:::{.version-solution}
```{r}
value <- 3
value %% 2 == 1

value <- 6
value %% 2 == 1
```
:::

**9. Write code to manipulate the vector of numbers `nums`. Your code should contain a `for()` loop that returns every other value in the vector, starting from the first element. That is, your code should return the 1st, 3rd, 5th, etc. values.**

::: callout-tip
You need to loop through the positions of `nums` (i.e., 1 through `length(nums)`) and find the odd positions. Use your ideas from Q8.
:::

:::{.version-solution}
```{r}
every_other <- c()
for (i in 1:length(nums)) {
  if (i %% 2 == 1) {
    every_other <- c(every_other, nums[i])
  }
}
every_other
```
:::

**10. Copy and edit your code from Q9. Your code should now return every other value in the vector, starting from the *second* element. That is, your code should return the 2nd, 4th, 6th, etc. values.**

:::{.version-solution}
```{r}
every_other <- c()
for (i in 1:length(nums)) {
  if (i %% 2 == 0) {
    every_other <- c(every_other, nums[i])
  }
}
every_other
```
:::

**11. Write a function called `every_other_num()`. This function should combine your code from Q9 and Q10 to return every other value in a vector. Include an argument called `start` that lets you choose whether you start from the first element (`start = 1`) or the second element (`start = 2`).** *Fill in the skeleton below.*

```{r}
# Complete the code below.
every_other_num <- function(){
  
  if(start == 1){

    
  } else if(start == 2) {

    
  }
  
}
```

:::{.version-solution}
```{r}
every_other_num <- function(vec, start){
  every_other <- c()
  if(start == 1){
    for (i in 1:length(vec)) {
      if (i %% 2 == 1) {
        every_other <- c(every_other, vec[i])
        }
      }
  } else if(start == 2) {
    for (i in 1:length(vec)) {
      if (i %% 2 == 0) {
        every_other <- c(every_other, vec[i])
        }
      }
  }
  return(every_other)
}
```
:::

**12. Test your function by running the code below.**

```{r}
# Don't change this code!
test <- c(1:10)

every_other_num(test, start = 1)
every_other_num(test, start = 2)
```

### Function 4: `shorten()`

**13. Write code to manipulate the vector of numbers `nums`. Your code should use a `while()` loop to find the first cumulative sum GREATER THAN 10.**

::: callout-tip
The $i^{th}$ cumulative sum of a vector is the sum of element $1$ through element $i$.

Consider the vector $(2,4,6,8,10)$. The first cumulative sum is $2$; the second cumulative sum is $2 + 4 = 6$; the third cumulative sum is $2 + 4 + 6 = 12$, etc.
:::

:::{.version-solution}
```{r}
total <- 0
ind <- 0
while(total <= 10) {
  ind <- ind + 1
  total <- total + nums[ind]
}
ind
```
:::

**14. Write a function called `shorten()`. This function should take in a vector, drop all values BEFORE the cumulative sum is greater than a provided number, and return the remaining values from the original vector. Adapt your `while()` loop from Q13 to do this.**

:::{.version-solution}
```{r}
shorten <- function(vec, limit) {
  total <- 0
  ind <- 0
  while(total <= limit) {
    ind <- ind + 1
    total <- total + vec[ind]
  }
  return(vec[-c(1:(ind-1))])
}
```
:::

**15. Write code to test your function.**

```{r}
# Write code to test the shorten() function here.
test <- 1:10
shorten(test, 2)
shorten(test, 20)
```

## Part 2: Find the Best Breakfast Burrito

**16. Once you have written your four functions correctly, run the following code:**

```{r}
my_vec <- c(39, 1.87, 48, 11, 8, 45, 21, 5, 12, 33, 9, 11, 108,
            4, 18, 5, 16, 17, 8, 48, 27, 24, 4, 22, 12, 27, 23,
            46, 42, 35, 15, 34, 36, 26, 18, 10, 18.21, 72.04,
            36.9, 41.81, 29, 89.75, 34.03, 20.18, 48.74, 15.76,
            31.86, 83.6, 43.55, 39.99, 23.55, 8.54, 24.71, 22.02,
            9.71, 62.14, 35.46, 16.61, 15.66, 21.29, 30.52,
            201.07, 45.81, 7.85, 30.13, 21.02, 22.62, 10.2, 6.02,
            30.12, 10.64, 31.72, 24.57, 14.43, 43.37, 89.93,
            44.72, 34.14, 13.62, 45.56, 22.96, 7.05, 29.99, 41.38,
            26.59, 23.04, 19.82, 50.73, 39.56, 8.88, 30.22, 85.85,
            5.78, 78.85, 29.52, 66.27, 44.06, 39.57, 24.43, 64.32,
            3.35, 67.45, 46.72, 48.44, 48.65, 49.1, 40.28, 19.04,
            13.6, 74, 29, 42.4, 6, 4, 49, 16.17, 22.62, 9.31,
            92.6, 9, 63.01, 30.12, 14, 31.72, 88.57, 15.3, 3.7,
            89.93, 72.4, 33.1, 29, 44.39, 56.29, 23.04, 1.99,
            51.15, 56.93, 13)
 # Has 138 elements! 

my_vec <- every_other_num(my_vec, start = 2)
# Should have 69 elements!

my_vec <- divide_and_round(my_vec)

my_vec <- every_other_num(my_vec, start = 1)
# Should have 35 elements!

my_vec <- shorten(my_vec, 350)
# Should have 20 elements!

my_vec <- my_vec[no_fours_or_fourteens(my_vec)]
# Should have 9 elements! 

my_vec
```

::: {.version-instructions}
::: callout-tip
### And the winner is...
If you have done everything correctly, your final vector will be nine numbers long. To find the name of a SLO establishment, turn these numbers into letters (where 1 = A, 2 = B, etc.).
:::
:::

:::{.version-solution}
Kreuzberg!
:::
