---
title: "Unit 1 Exam"
format: 
  html:
    fig-width: 7
    fig-height: 5
filters:
- flourish
df-print: kable
theme: flatly
execute:
  echo: true
  message: false
  warning: false
  error: true
---

<!-- create a vector with c() -->
<!-- create a matrix, data frame, list -->
<!-- max, min, mean -->
<!-- indexing -->
<!-- relational comparisions -->
<!-- adding logical values -->

Benito Antonio Martínez Ocasio, known professionally as Bad Bunny, is a Puerto
Rican rapper, actor, singer-songwriter, and record producer. Bad Bunny has
relased six studio albums since 2018 with multiple obtaining platnum status. 

We've obtained information on each album, specifically what year it was 
released, the top chart position it reached in the US, how many sales the 
album had, and the number of tracks in the album. 

```{r}
#| label: album-info-numeric
#| echo: true

year_released <- c(2018, 2020,  2020, 2022, 2023, 2025)
top_us_chart_position <- c(11, 2, 1, 1, 1, 1)
sales <- c(66000, 1444000, 348000, 3400000, 184000, 2833000)
no_tracks <- c(15, 20, 16, 23, 22, 17)

```

**1. We would like to combine these four pieces of information into a _single_
object. Complete the code below to combine these vectors into a matrix.**

```{r}
#| flourish:
#| - target: 
#|     - '   ncol'
#|     - mask: true
#|     - style: "text-decoration: underline;"
#| eval: false

matrix(
  c(year_released, top_us_chart_position, sales, no_tracks), 
     ncol = 4
  )
```
**2. Will this code result in any errors? If yes, what error would be printed? 
If no, why will there be no errors?**

**3. Next we would like to add the album's name and the top award it received. 
Can we add these vectors as columns in the matrix we made in #1? If yes, how 
would this be done? If no, why can this not be done?**

```{r}
#| label: album-info-character
#| echo: true

albums <- c("X 100pre", 
            "YHLQMDLG", 
            "El Último Tour Del Mundo",
            "Un Verano Sin Ti", 
            "Nadie Sabe Lo Que Va a Pasar Mañana", 
            "Debí Tirar Más Fotos")
  
awards <- c(NA, 
            "Best Música Urbana Album, Latin Grammys",
            "Best Urban Music Album, Grammys", 
            "Top Latin Album, Billboard Music Awards",
            "Top Latin Album, Billboard Music Awards",
            "Album of the Year, Grammys"
            )
```

**4. Finally, we are interested in incorporating the tracklist for each album. 
What structure of object (i.e., vector, dataframe, list, matrix) would we 
need to use to incorporate this information into the existing album
information?**

::: columns
::: {.column width="48%"}
```{r}
#| label: album-tracks-1

X <- c("Ni Bien Ni Mal", 
       "200 MPH", 
       "¿Quién Tú Eres?", 
       "Caro", 
       "Tenemos Que Hablar", 
       "Otra Noche en Miami", 
       "Ser Bichote", 
       "Si Estuviésemos Juntos", 
       "Solo de Mí", 
       "Cuando Perriabas", 
       "La Romana", 
       "Como Antes", 
       "RLNDT", 
       "Estamos Bien", 
       "Mía")
```
:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
```{r}
#| label: album-tracks-2

YHLQMDLG <- c("Si Veo a Tu Mamá", 
              "La Difícil", 
              "Pero Ya No", 
              "La Santa",
              "Yo Perreo Sola", 
              "Bichiyal", 
              "Soliá", 
              "La Zona", 
              "Que Malo", 
              "Vete", 
              "Ignorantes", 
              "A Tu Merced", 
              "Una Vez", 
              "Safaera", 
              "25/8", 
              "Está Cabrón Ser Yo",
              "Puesto Pa' Guerrial", 
              "P FKN R", 
              "Hablamos Mañana", 
              "<3")
    
```
:::
:::


<!-- extension question: add a list element to a list -->

