---
title: "Week 2: R is Friend"
number-sections: true
number-depth: 2
---

```{r}
#| include: false
#| label: packages-setup

library(tidyverse)
```

This week we're going to learn more about the building blocks of the R 
programming language. 

In this week's reading, we have:

::: columns
::: {.column width="48%"}
#### 📖 Readings: 60 min

#### 🏋🏽‍♀️️ Activities: 30 min

:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
#### 📽 Videos: 0 min

#### 💻 Tutorials: 0 min
:::
:::

------------------------------------------------------------------------

</br>

## R is Friend

::: callout
## Learning Objectives

+ Identify three of the foundational data types in R
+ Explain similarities and differences between different object types
+ Use indexing to extract elements from different object types
+ Use functions to calculate summary statistics
+ Use logical values and indexing to calculate summary statistics
:::

### 📖 Required Reading: [R is Friend](
https://r-is-your-friend.github.io/course-pack/02-r-is-friend.html)

There are **a lot** of check-in questions this week! Each section contains 1-3 
questions targeting the ideas covered in that section. We encourage you to 
keep track of your answers as you progress!

::: {.callout-check-in}

For each of the examples below, identify what data type do you believe the
vector will have.

1. `c(67, 13, 1989, TRUE)`
2. `c("hello", "goodbye", TRUE)`
3. `c("TRUE", FALSE, TRUE, "FALSE")`

4. Draw out how this matrix will look:

```{r}
#| eval: false
#| label: matrix-draw-dimensions

shows <- c("Tale of a Thousand Stars", 
           "Bad Buddy", 
           "Moonlight Chicken", 
           "Not Me")

matrix(shows, ncol = 2)
```

```{r}
#| echo: false
#| label: identify-object-fruit-info

fruits <- c("apple",
            "orange", 
            "lemon", 
            "lime", 
            "mango", 
            "peach")

trees <- c("Malus domestica", 
           "Citrus sinensis", 
           "Citrus limon", 
           "Citrus aurantifolia", 
           "Mangifera indica", 
           "Prunus persica")

harvest_window <- c(3, 5, 12, 12, 5, 3)
```

5. What type of object is this?

```{r}
#| label: identify-object-1
#| echo: false
 
list(trees, harvest_window)
```

6. What type of object is this?

```{r}
#| label: identify-object-2
#| echo: false

matrix(harvest_window, nrow = 3)
```

7. What type of object is this?

```{r}
#| label: identify-object-3
#| echo: false

data.frame(fruits, trees)
```

8. What type of object is this?

```{r}
#| label: identify-object-4
#| echo: false

list(tree = trees, harvest_window = harvest_window)
```

9. What type of object is this?

```{r}
#| label: identify-object-5
#| echo: false

trees
```

```{r}
#| label: vector-check-in
fruits <- c("apple",
            "orange", 
            "lemon", 
            "lime", 
            "mango", 
            "peach")
```

10. Grab the first three elements of the vector named `fruits`.

11. Grab the first, third, and fifth element of the vector named `fruits`.

```{r}
trees <- c("Malus domestica", 
           "Citrus sinensis", 
           "Citrus limon", 
           "Citrus aurantifolia", 
           "Mangifera indica", 
           "Prunus persica")

tree_matrix <- matrix(trees, ncol = 2)
tree_matrix
```

12. Grab the tree in the second row, second column of the matrix named 
`tree_matrix`.

13. Grab the trees in the first two rows of the matrix named `tree_matrix`. 

14. Grab the trees in the second column of the matrix named `tree_matrix`.

```{r}
harvest_window <- c(3, 5, 12, 12, 5, 3)

fruit_tree_harvest <- list(fruit = fruits, 
                           tree = trees, 
                           harvest_window = harvest_window)
fruit_tree_harvest
```

15. Grab the first component (`fruit`) of the list named `fruit_tree_harvest`.  

16. Grab the third tree from the `tree` component of the list named  
`fruit_tree_harvest`.

```{r}
fruit_trees <- data.frame(fruits, trees, harvest_window)
fruit_trees
```

17. Grab the second column (`trees`) of the `fruit_trees` dataframe.

18. Grab the second and third columns (`trees`, `harvest_window`) of the
`fruit_trees` dataframe.

19. Grab the second row of all three columns of the `fruit_trees` dataframe.

20. What will be the output of the following operation?

```{r}
#| eval: false
#| label: add-two-vectors-exercise

vec1 <- c(1, 2, 3, 4)
vec2 <- c(5, 6, 7, 8)

vec1 + vec2
```

(a) `1 2 3 4 5 6 7 8`
(b) `6 8 10 12`
(c) this code will have an error
:::

