---
title: "Week 3: Control Flow"
number-sections: true
number-depth: 2
---

```{r}
#| include: false
#| label: packages-setup

library(tidyverse)
```

This week we're going to learn about control flow---statements in a computer
program that that determines when, how, and how many times code is evaluated. 

In this week's reading, we have:

::: columns
::: {.column width="48%"}
#### 📖 Readings: 60 min

#### 🏋🏾‍♂️ Activities: 30 min

:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
#### 📽 Videos: 0 min 

#### 💻 Tutorials: 0 min
:::
:::

## Control Flow

::: callout
## Learning Objectives

- What a conditional statement is
- How to create a simple conditional statement with `if ()` and `else`
- How to series of conditional statements with `else if ()`
- What it means for a function to be vectorized
- How to make a vectorized conditional statement
- Where to access help on an R function
- Recognize when to use a `for ()` loop instead of a `while ()` loop
- Understand how to index a loop
:::

### 📖 Required Reading: [Control Flow](
https://r-is-your-friend.github.io/course-pack/03-control-flow.html)

By the end of the reading, you should be able to answer the following questions:

::: {.callout-check-in}
1. Which of the following is the correct way to format a conditional statement?

(a)
```{r}
#| label: question-1-a
#| eval: false

if (num < 100){
  print(num)
}
else {
  print("banana")
}
```

(b)
```{r}
#| label: question-1-b
#| eval: false

if (num < 100){
  print(num)
} else {
  print("banana")
}
```

(c)
```{r}
#| label: question-1-c
#| eval: false

if (num < 100){ print(num) } else { print("banana") }
```

(d)
```{r}
#| label: question-1-d
#| eval: false

if (num < 100) print(num) else print("banana") 
```

2. With `if` + `else` you have ____ condition(s) to check and ____ possible 
outcome(s).

3. In the IRS tax brackets we had ____ possible outcomes and ____ conditions that needed to be checked. 

4. In general, if you have $n$ possible outcomes, you will **always** have
____ conditions to check. 

5. A function is "vectorized" if it accepts a 
[vector / matrix / dataframe / list] with length 
[equal to / greater than / less than] 1 as an input. 

6. Which of the following options would open the help file for the `mean()`
function? Select all that apply!

(a) Type `mean` in the Console
(b) Type `?mean` in the Console
(c) Search for mean in the search bar of the Packages tab
(d) Search for mean in the search bar of the Help tab

7. The `.unmatched` argument of `case_when()` throws an error when ____. 

8. The `.default` argument of `case_when()` sets a default value for inputs that
____. 

9. True or false, you can use **both** the `.unmatched` and `.default`
arguments. Meaning you can set `.unmatched =  "error"` **and** 
`.default = "zero"`?

10. Match the loop to its correct setting. 

::: columns
::: {.column width="48%"}

`for()` Loop

</br>

`while()` Loop
:::

::: {.column width="4%"}
:::

::: {.column width="48%"}

Runs until a condition is not met (`FALSE`)

Runs a set number of times
:::
:::
