---
title: "Week 3 Part 2: Writing Simple Functions in R"
number-sections: true
number-depth: 2
---

```{r}
#| include: false
#| label: packages-setup

library(tidyverse)
```

At this point, you've encountered almost 20 functions either built-in to R 
(e.g., `sqrt()`) or that live in specific packages (e.g., `case_when()`). For
each function, we've learned how we can use it to get the results we were after,
but we haven't dug into what makes a function a function. For the second half
of this week, we're going to learn about what makes a function a function.

In this week's reading, we have:

::: columns
::: {.column width="48%"}
#### 📖 Readings: 60 min

#### 🏋🏾‍♂️ Activities: 30 min

:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
#### 📽 Videos: 0 min 

#### 💻 Tutorials: 0 min
:::
:::

------------------------------------------------------------------------

</br>

## Writing Functions in R

::: callout
## Learning Objectives

- Understand the components of a function
- Outline how function arguments are used
- Contrast arguments with and without default values
- Illustrate how control flow can be used inside the body of a function
- Explain the differences between explicit and implicit function returns
:::

### 📖 Required Reading: [Writing Simple Functions in R](
https://r-is-your-friend.github.io/course-pack/03b-writing-functions.html)

By the end of the reading, you should be able to answer the following questions:

::: {.callout-check-in}
1.  Now that we have created the `rescale()` function, let's check our intuition
for what the function should output (return) for different inputs.

```{r}
x <- 3
y <- 10:20
z <- c(1, 4, 5, 6, "dog")
```

a.  What would you expect `rescale()` to output when `x` is input?
b.  What would you expect `rescale()` to output when `y` is input?
c.  What would you expect `rescale()` to output when `z` is input?

We know `rescale()` works, so let's test our intuition for what would happen
for the following scenarios. 

2. `rescale(nums, method = "normalize")`
3. `rescale(nums, method = "min-max")`
4. `rescale(nums, na.rm = TRUE)`

We've incorporated a input check for the size of `x`, but not for the size of
`na.rm` or `method`. 

5. What `length()` should we expect `na.rm` to have?

6. What `length()` should we expect `method` to have?

7. What should the final set of input checks be?
:::
