---
title: "Week 3 Part 2: Writing Simple Functions in R"
number-sections: true
number-depth: 2
filters:
 - flourish
---

```{r}
#| include: false
#| label: packages-setup

library(tidyverse)
```

At this point, you've encountered almost 20 functions either built-in to R 
(e.g., `sqrt()`) or that live in specific packages (e.g., `case_when()`). For
each function, we've learned how we can use it to get the results we were after,
but we haven't dug into what makes a function a function. For the second half
of this week, we're going to learn about what makes a function a function.

In this week's reading, we have:

::: columns
::: {.column width="48%"}
#### 📖 Readings: 60 min

#### 🏋🏽‍♀️ Activities: 30 min

:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
#### 📽 Videos: 0 min 

#### 💻 Tutorials: 0 min
:::
:::

------------------------------------------------------------------------

</br>

## Writing Functions in R

::: callout
## Learning Objectives

- Understand the components of a function
- Outline how function arguments are used
- Contrast arguments with and without default values
- Illustrate how control flow can be used inside the body of a function
- Explain the differences between explicit and implicit function returns
:::

### 📖 Required Reading: [Writing Simple Functions in R](
https://r-is-your-friend.github.io/course-pack/03b-writing-functions.html)

By the end of the reading, you should be able to answer the following questions:

::: {.callout-check-in}
Suppose `divide()` is a function with three arguments: 

1. `x`, a number
2. `y`, a number
3. `round`, a logical (`TRUE/FALSE`) value

The function calculates `x/y` and then decides whether or not to round the
result (depending on whether `round = TRUE` or `round = FALSE`). We want to
calculate `17/3` and then round the result. For each of the *function calls*
below, indicate if this would result in

(a) An error
(b) No error, but an incorrect result
(c) The correct result

1. `divide(17, 3, TRUE)`
2. `divide(round = TRUE, y = 3, x = 17)`
3. `divide(17, 3, round = TRUE)`
4. `divide(round = TRUE, 3, 17)`
5. `divide(TRUE, 17, 3)`

Now that we have created the `rescale()` function, let's check our intuition
for what the function should output (return) for different inputs.

6.  What would you expect `rescale()` to output when `test_single` is input?
7.  What would you expect `rescale()` to output when `test_missing` is input?
8.  What would you expect `rescale()` to output when `test_type` is input?

Let's think about what you would expect the `find_tax_rate()` function to do 
for different scenarios.

9.  What would you expect `find_tax_rate()` to output when a *character* object
is input?  Why?
10.  What would you expect `find_tax_rate()` to output when no arguments are
provided? Why?
11.  What would you expect `find_tax_rate()` to output when a numeric **vector**
is input? Why?


Suppose we want to use a *loop* to rescale the vectors `cpi`, `ppe`, 
and `pci`.  What should go in the blanks below?


```{r}
#| eval: false
#| flourish:
#| - target: 
#|     - "blah"
#|     - mask: true
#|     - style: "text-decoration: underline;"
#| - target-rx: 
#|     - "blahblah"
#|     - mask: true
#|     - style: "text-decoration: underline;"


economy_data <- list(cpi, ppi, pce)

for(i in blah) {                                       
  
  economy_data[[i]] <- rescale(economy_data[blahblah])     
  
}
```

::: columns
::: {.column width="48%"}
Suppose we've created the following function:

```{r}

friend_or_foe <- function(){
  if_else(pet == "cat", 
          "friend", 
          "foe")
}
```
:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
And we have the following value in our global environment:

```{r}
pet <- "cat"
```

:::
:::

13. What would `friend_or_foe()` output?

14. What if we ran the following code `pet <- "parrot"`? What would 
`friend_or_foe()` output?

15. What if there was no object named `pet` in the global environment? What 
would `friend_or_foe()` output?
:::
