---
title: "Lab 2: R is Friend"
format: 
  html:
    toc: true      

author: SOLUTIONS
error: true
eval: true
---

<!-- We need to decide what our AI policy is! -->

## Create an R Script

Last week, you downloaded a template R script, but you've learned how to make
one of these from scratch. So, this week you are going to make your own R script
for Lab 2!

1. Create a new R script, give it a descriptive and appropriate file name, and
save it in your Week 2 subfolder (of your STAT 1810 folder). 

2. In the newly created R script, add a comment to the top that describes the
contents of this document. *Hint: it's the code for Lab 2!*

The top of your R Script should have a code comment (marked by a `#`) followed
by a few words describing the content of your script. Something like:

```{r}
# Code for Lab 2
```

## Pismo Beach Water Temperature

It's Saturday, you have some free time, and you're trying to decide if it's warm
enough to go swimming at Pismo Beach! You look up the water temperature at Pismo
Pier on a surf watch website, but the temperature is only listed in degrees
Celsius! What does that even mean???

3. Create a new object to hold to current water temperature in degrees
Celsius -- it's listed as 11 degrees. Make sure you use a **descriptive** 
name for the object!

```{r}
temp_C <- 11
```

4. Write one line of code to turn degrees Celsius into degrees Fahrenheit,
incorporating the variable you create in #3. What is the water temperature?

```{r}
(temp_C * (9/5)) + 32
```

Since it was too cold to swim on Saturday, you decided to keep track of the
water temperature to see what days might have temperatures warm enough for 
swimming. You observed the following temperatures:

- Sunday: 12°C
- Monday: 13°C
- Tuesday: 16°C
- Wednesday: 14°C
- Thursday: 17°C
- Friday: 16°C
- Saturday: 18°C

5. Create an object to store these seven temperatures.  Make sure you use a
**descriptive** name for the object!

```{r}
temperatures <- c(12, 
                  13, 
                  16, 
                  14, 
                  17, 
                  16, 
                  18)
```

6. Create an object to store the seven days these temperatures are associated 
with.  Make sure you use a **descriptive** name for the object!

```{r}
days <- c("Sunday", 
          "Monday", 
          "Tuesday", 
          "Wednesday", 
          "Thursday", 
          "Friday", 
          "Saturday")
```

7. Now that we have temperatures **and** their associated days of the week, 
let's combine these into **one** object. Create an object that stores *both* the
temperatures you made in #5 **and** the days you made in #6. 

```{r}
days_temps <- data.frame(days = days, 
                         temps = temperatures)
```

8. Now, let's change the temperature values from Celsius into Fahrenheit. Be
sure to save the updated values back into the original object (like you did 
in the Practice Activity)! *Hint: You should borrow the mathematical steps you used in #4!*

```{r}
days_temps$temps <- (days_temps$temps * (9/5)) + 32
```

9. If 60 degrees Fahrenheit is your threshold for swimming in the ocean, how
many days had "swimable" temperatures?

```{r}
sum(days_temps$temps >= 60)
```

## Do Other People Like Your Favorite Movies?

Let's use [Rotten Tomatoes](https://www.rottentomatoes.com/) to find the average
rating for five of your favorite movies. Do other people like them as much as
you do?

10. Create a variable that contains the Rotten Tomatoes ratings for your five
favorite movies. Make sure you use a **descriptive** variable name.

```{r}
top5_ratings <- c(86, 
                  44, 
                  94, 
                  70, 
                  91)
```

11. Create a variable that contains the names of your five favorite movies. 
Remember the names should go in the same order as the ratings! Make sure you
use a **descriptive** variable name.

```{r}
top5_movies <- c("Never Been Kissed",
                 "Fire Island",
                 "Singing in the Rain", 
                 "In the Mood for Love", 
                 "Sound of Music")
```

12. Write a line of code to find the maximum rating for your top five movies. 
Write a second line of code to find the minimum rating.

```{r}
min(top5_ratings)
max(top5_ratings)
```

13. Write a line of code to find the average rating for your top five movies.

```{r}
mean(top5_ratings)
```

We've asked some of the statistics faculty what their favorite movie is and
assembled them into a vector. 

```{r}
stats_prof_movies <- c("10 Things I Hate About You", 
                       "The Notebook", 
                       "Legally Blonde", 
                       "Eternal Sunshine of the Spotless Mind", 
                       "Mean Girls")
```

14. Use the `%in%` operator to check if any of these movies are included in 
the top five movies you made in #11. Here is an example of how to use the
`%in%` operator:

```{r}
"bananna" %in% c("apple", "kiwi", "banana", "orange")
```

```{r}
stats_prof_movies %in% top5_movies
```

## Reviewing Your Work

15. What variables have you created during this lab? Using a code comment, list
all of the variable names.

```{r}
# I've created 7 variables: 
# temp_C, temperatures, days, days_temps, top5_ratings, top5_movies, stats_prof_movies
```

## Make Your Code Readable

16. Add comments throughout your script that describe what your code is doing. 
