---
title: "PA 12: Find the mystery animal"
filters:
  - versions
versions:
  - student:
      out-dir: ".."
      yaml:
        author: "Your Name"
        embed-resources: true
  - instructions:
      out-dir: ".."
      yaml:
        eval: false
  - solution:
      out-dir: ".."
      out-file: "pa12-solution.qmd"
      yaml:
        author: "SOLUTIONS"
        embed-resources: true
format: 
  html:
    code-tools: true
    toc: true      
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

A group of scientists is studying the diet of a particular mystery animal. They
have performed an experiment: First, they recorded the weights of each individual
animal in their sample.  Then, they fed the animals only roasted duck for one
year, after which they recorded the new weights of the animals.

The dataset below contains the weights *before* and *after* a year 
of the roasted duck diet.

:::{.version-instructions}
[Download starter .qmd file](../student/pa12-regression.qmd)

[Download data -- `12-mystery_animal.csv`](../../data/12-mystery_animal.csv)
:::

```{r}
#| versions: solution

library(tidyverse)

dat <- read_csv(here::here("data", "12-mystery_animal.csv"))
```

## Data Exploration

1. Make a scatter plot of this data, with the `weight_before` measurements on the x-axis 
and the `weight_after` measurements on the y-axis.  

2. What appears to be the slope? Why does this make sense in the context of this data?

::: {.versions-solution}

Both smaller and bigger animals might gain some weight, but the overall weights
don't change much - we expect `weight_after` to be very similar to `weight_before`.

:::

3. Why is it difficult to estimate the intercept from this data?

::: {.versions-solution}

The data starts at around 100, since the animals weigh between 100 and 400 pounds.
So, we can't see what value the y-axis would have at x = 0.

:::

## Model calculation

1. Write a function called `linear_model` that takes an `x` and `y` vectors as
input, and returns the calculated slope and intercept.

```{r}
#| version: solution

linear_model <- function(x, y) {
  
  slope <- cor(x, y)*sd(y)/sd(x)
  
  int <- mean(y) - slope*mean(x)
  
  return(list(slope = slope, intercept = int))
  
}
```



2. Use your new function to *fit* a linear model to this data, predicting the
`after` weights from the `before` weights.

```{r}
#| version: solution

mod <- linear_model(x = dat$weight_before, y = dat$weight_after)
mod
```


2. What is the *slope* of this model?  Does that match what you predicted? Why
do you think it's not *exactly* equal to 1?

::: {.versions-solution}
1.19.  Yes, this is close to 1.  But there is always some variability in how
much weight the animals gain, so the relationship is not exact for this particular
dataset.
:::


3. What is the *intercept* of this model?  What does that suggest to you about
the Duck Diet?  Did the animals gain weight, lose weight, or neither?

::: {.versions-solution}

2.85.  This means they gained a little weight; about 3 pounds on average.

:::

## Predictions and residuals

4. Add a column to the dataset named `pred_after`, which contains the *predicted*
weight of the animals after a year of duck diet, based on this model.

```{r}
#| versions: solution

dat <- dat |>
  mutate(
    pred_after = 1.19*weight_before + 2.85
  )
```



5. Add a column to the dataset named `residual`, equal to the *predicted* value
subtracted from each *actual observed* value.


```{r}
#| versions: solution

dat <- dat |>
  mutate(
    resid = weight_after - pred_after
  )
```
 

6. Make a scatterplot with the residuals on the y-axis and the predicted values
on the x-axis.  

```{r}
#| versions: solution

dat |>
  ggplot() +
  aes(x = pred_after, y = resid) +
  geom_point()
```



7. What is the mystery animal that was fed the roasted duck diet?

::: {.versions-solution}

A Shark!

:::
