---
title: "PA 8: Twinning"
subtitle: "Data Joins"
filters:
  - versions
versions:
  - student:
      out-dir: ".."
      yaml:
        author: "Your Name"
        embed-resources: true
  - instructions:
      out-dir: ".."
      yaml:
        eval: false
  - solution:
      out-dir: ".."
      out-file: "pa8-solution.qmd"
      yaml:
        author: "SOLUTIONS"
        embed-resources: true
format: 
  html:
    code-tools: true
    toc: true  
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

```{r}
#| label: setup

# load packages
library(readxl) 
library(tidyverse)
library(knitr)
```

:::{.version-instructions}

[Download starter .qmd file](../student/pa8-joins.qmd)


Today we will be exploring "twin towns" (also called sister cities). These are cities that have some sort of social or political relationship with each other, usually for the purpose of cultural exchange. A list of twin towns is kept up on Wikipedia, which is the source of our data today!

The data come to us from [TidyTuesday](https://github.com/rfordatascience/tidytuesday/blob/main/data/2026/2026-05-12/readme.md#data-dictionary) -- originally compiled by 
[Ahmad Barclay](https://github.com/bothness/twin-cities) and further curated by [Nicola Rennie](https://github.com/nrennie) for TidyTuesday.

We also created a dataset `dists` with the following columns:

- `source`: same as in `links`
- `target`: same as in `links`
- `dist`: distance between paired cities in km

:::

```{r}
cities <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-05-12/cities.csv')

links <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-05-12/links.csv')

dists <- read_csv('https://raw.githubusercontent.com/r-is-your-friend/course-assessments/main/data/08-dists.csv')
```

1. Check out the data. What is one observation in the `cities` dataset and what kind of information do we get from this data? What is one observation in the `links` dataset and what kind of information do we get from this data?






The `links` data contains all of the pairs of cities, but we can't learn anything interesting just from the city IDs! We want to add the city information on.

2. What are the key(s) between the `city` and `links` data?


As the data dictionary states, the order of "source" and "target" city is arbitrary. 

3. Created a new dataframe called  `linked_cities` that includes **all** of the information in the `cities` data for both cities in each pair of twinned cities.

Note: This means you will have *two* name, latitude, longitude, etc. columns for each row. Think about how you might want to name columns to keep things from being confusing!

```{r}
#| version: student

# Q3 code
```


```{r}
#| version: solution

linked_cities <- links |> 
  left_join(cities, join_by(source == id)) |> 
  rename_with(.fn = \(x) paste0(x, "_1"), 
              .cols = name:continent) |> 
  left_join(cities, join_by(target == id)) |> 
  rename_with(.fn = \(x) paste0(x, "_2"), 
              .cols = name:continent)
```

4. Are any of the twinned cities missing city information like their name?

```{r}
#| version: student

# Q4 code
```


```{r}
#| version: solution

linked_cities |> 
  filter(is.na(name_1) | is.na(name_2))
```

5. Q4 implies that we *could* have used which of the following joins (or all of them)? inner, left, right, and/or full?

:::{.version-solution}

We didn't find that any of the twinned cities were missing information from the `cities` data, keeping all of the `linked_cities` observations, so we definitely could have used inner or left join with the same outcome, (assuming `linked_cities` was the left dataset). If you did some more investigating, you would find that the `cities` data does not include any cities that are not in the `linked_cities` data as well. Since the same exact keys are present in both datasets, any of the joins would have resulted in the same outcome here!

:::


6.  Join the distances between cities on to your `linked_cities` data. Pay close attention: what identifies one observation in the `dists` data? Your `linked_cities` data should still have 10,596 observations.

```{r}
#| version: student

# Q6 code
```

```{r}
#| version: solution

linked_cities <- linked_cities |> 
  left_join(dists, 
            by = join_by(source, target))
```


7. Which two twin cities are the *farthest* apart? Output one row with this answer, including the two cities names, countries, and the distance between them.

```{r}
#| version: student

# Q7 code
```

```{r}
#| version: solution

linked_cities |> 
  slice_max(order_by = dist,
            n = 1) |> 
  select(starts_with("name"),
         starts_with("country_"),
         dist) |> 
  kable()
```


8. Which two twin cities are the *closest* together? Is this a mistake in the data!? Look up the two cities on a map.

```{r}
#| version: student

# Q8 code
```


```{r}
#| version: solution

linked_cities |> 
  slice_min(order_by = dist,
            n = 1) |> 
  select(starts_with("name"),
         starts_with("country_"),
         dist) |> 
  kable()
```

9. Which country has the most twinned cities *within* the same country?

```{r}
#| version: student

# Q9 code
```


```{r}
#| version: solution

linked_cities |> 
  filter(country_1 == country_2) |> 
  group_by(country_1) |> 
  count() |> 
  ungroup() |> 
  slice_max(order_by = n) |> 
  kable()
```


10. CHALLENGE: Which continent has the most twinned cities with North America? Create a table that has each continent and the number of cities that they have that are twinned with a city in North America.

```{r}
#| version: student

# Q10 code
```


```{r}
#| version: solution

linked_cities |> 
  filter(continent_1 == "North America"|
           continent_2 == "North America") |> 
  group_by(continent_1, continent_2) |> 
  count() |> 
  pivot_longer(starts_with("continent"),
               names_to = "order",
               values_to = "continent") |> 
  filter(continent != "North America") |> 
  summarize(n = sum(n), .by = continent) |> 
  arrange(desc(n)) |> 
  kable()
```


:::callout-note
## Canvas Submission

Your answers to Q7-9 will be your answers to the Canvas Quiz!

:::
