---
title: "PA 9: Cancel Culture"
subtitle: "Strings and Date Variables"
filters:
  - versions
versions:
  - student:
      out-dir: ".."
      yaml:
        author: "Your Name"
        embed-resources: true
  - instructions:
      out-dir: ".."
      yaml:
        eval: false
  - solution:
      out-dir: ".."
      out-file: "pa9-solution.qmd"
      yaml:
        author: "SOLUTIONS"
        embed-resources: true
format: 
  html:
    code-tools: true
    toc: true      
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

```{r setup}
# load packages
library(tidyverse)
```


The data are compiled by Grant Witness.


You can find the data dictionary [here](https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-05-06/readme.md#data-dictionary)

```{r}
nsf_terminations <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-06/nsf_terminations.csv')

```

## Data Context

First, as always, let's spend some time getting acquainted with the data. 

1. How does this data relate to you and your life? What are your impressions of the social and political context of this data? What questions might you want to answer using this data?


2. Looking at the data and the data dictionary together, do you have any questions about what any of the columns represent? 


3. What does one row of the data represent? What is included in the sample -- are any possible observations in a full population left out?


4. Which variables in the data represent dates? Are they stored as dates in the `nsf_terminations` data, or do we need to convert them to dates? If you need to convert them, do that.


## Data Cleaning

Next, let's clean up the data a little bit and adding some variables that we will use later.

Implement the following in one dplyr pipeline. But go one step at a time to catch any bugs as you go!

5. The city name is annoyingly mostly in all caps but sometimes in title case -- edit the `org_city` variable so that it is all in title case.

6. Create two new variables that give the year that the grant started and the intended end year of the grant. Call these variables `start_year` and `expected_end_year`.

7. Create a variable called `diversity` that gives the *number of times* that the word "diversity" or "equity" appears in the grant abstract. Be careful - make sure this does not depend on case!

```{r}
#| version: solution

nsf_terminations <- nsf_terminations |> 
  mutate(org_city = str_to_title(org_city),
         start_year = year(nsf_startdate),
         expected_end_year = year(nsf_expected_end_date),
         diversity = str_count(str_to_lower(abstract), 
                               "diversity|equity")
         )
```


## Find that Grant




String things:

- `str_length()` on title
- `str_detect()` to find certain types of grants

Date things:
- Create a variable for the day of the week that termination letters were sent
