---
title: "PA 9: Cancel Culture"
subtitle: "Strings and Date Variables"
filters:
  - versions
versions:
  - student:
      out-dir: ".."
      yaml:
        author: "Your Name"
        embed-resources: true
  - instructions:
      out-dir: ".."
      yaml:
        eval: false
  - solution:
      out-dir: ".."
      out-file: "pa9-solution.qmd"
      yaml:
        author: "SOLUTIONS"
        embed-resources: true
format: 
  html:
    code-tools: true
    toc: true      
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

```{r setup}
# load packages
library(tidyverse)
```


The data are compiled by Grant Witness.

There is both a Tidy Tuesday version, and an updated version I can download from Grant Witness... Maybe we stick with tidy tuesday since it has the cruz list column which is fun? There are 2000 observations in the updated data though. It also doesn't have a clean data dictionary.

You can find the data dictionary [here](https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-05-06/readme.md#data-dictionary)

```{r}
nsf_terminations <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-06/nsf_terminations.csv')

```


String things:

- org_city to title case `str_to_title()`
- count number of times the string "diversity" (or something else) is in the abstract `str_count()`
- "clean up" the grant name by replacing words `str_replace_all()`
- `str_count()` on title
- `str_detect()` to find certain types of grants

Date things:
- Create year start and year end variables
- Create a variable for the day of the week that termination letters were sent
- Calculate the total length of the grant in days
- Calculate the number of days that had passed for the grant before it was terminated
- Calculate the % of the grant time that was allowed
