---
title: "PA 9: Cancel Culture"
subtitle: "Strings and Date Variables"
filters:
  - versions
versions:
  - student:
      out-dir: ".."
      yaml:
        author: "Your Name"
        embed-resources: true
  - instructions:
      out-dir: ".."
      yaml:
        eval: false
  - solution:
      out-dir: ".."
      out-file: "pa9-solution.qmd"
      yaml:
        author: "SOLUTIONS"
        embed-resources: true
format: 
  html:
    code-tools: true
    toc: true      
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

```{r setup}
# load packages
library(tidyverse)
library(knitr)
```

[[NEED MORE PREAMBLE]]

The data are compiled by Grant Witness.


You can find the data dictionary [here](https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-05-06/readme.md#data-dictionary)

```{r}
nsf_terminations <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-06/nsf_terminations.csv')

```

## Data Context

First, as always, let's spend some time getting acquainted with the data. 

1. How does this data relate to you and your life? What are your impressions of the social and political context of this data? What questions might you want to answer using this data?


:::{.version-student}

*Answer in markdown here after discussing with your group.*

:::

:::{.version-solution}

*TBD*

:::


2. Looking at the data and the data dictionary together, do you have any questions about what any of the columns represent? 

:::{.version-student}

*Answer in markdown here after discussing with your group.*

:::

:::{.version-solution}

*Answers will vary -- I was somewhat unsure of what the `usaspending_obligated` variable meant at first.*

:::

3. What does one row of the data represent? What is included in the sample -- are any possible observations in a full population left out?

:::{.version-student}

*Answer in markdown here after discussing with your group.*

:::

:::{.version-solution}

*Each row represents a grant that was terminated between April 8, 2025 and May 6th, 2025 (when the data was posted). The data is curated by "crowd-sourcing from researchers and program administrators," so it is possible that it is not a complete set. In addition, as of July 2026, these issues are ongoing, so this data does not include all grants that were terminated after May 6th 2025, nor information on grants that were reinstated out of this list.*

:::

4. Which variables in the data represent dates? Are they stored as dates in the `nsf_terminations` data, or do we need to convert them to dates? What is the format? If you need to convert them, do that.

:::{.version-student}

*Answer in markdown here after discussing with your group.*

:::

:::{.version-solution}

*`termination_letter_date`, `nsf_startdate` and `nsf_expected_end_date` are all dates in the form "YYYY-MM-DD" (or "ymd"). They are already read in as dates, so we don't need to do anything!*

:::

## Data Cleaning

Next, let's clean up the data a little bit and adding some variables that we will use later.

Implement the following in one dplyr pipeline. But go one step at a time to catch any bugs as you go!

5. The city name is annoyingly mostly in all caps but sometimes in title case -- edit the `org_city` variable so that it is all in title case.

6. Create two new variables that give the year that the grant started and the intended end year of the grant. Call these variables `start_year` and `expected_end_year`.

7. Create a variable called `diversity` that gives the *number of times* that the word "diversity" or "equity" appears in the grant abstract. Be careful - make sure this does not depend on case!

```{r}
#| version: student

# Q5-7 data cleaning code

```



```{r}
#| version: solution

nsf_terminations <- nsf_terminations |> 
  mutate(org_city = str_to_title(org_city),
         start_year = year(nsf_startdate),
         expected_end_year = year(nsf_expected_end_date),
         diversity = str_count(str_to_lower(abstract), 
                               "diversity|equity"),
         title_len = str_length(project_title))
```


## Find that Grant

Later this week, we will explore general trends in the cancelled grants more. However, each grant tells a whole story and represents many people and resources that were impacted by the termination.

Follow the clues below to narrow our investigation down to one grant.

:::{.version-instructions}

<!--Add more context around why we are looking at one grant?? There is really no reason to search for this grant in particular... I just thought it was interesting and worked with some relevant skills to narrow it down -->

:::

8. The grant number starts with "23". Filter to these grants using one function from `stringr`. (Hint: do you need to change the variable type to do this?). 

```{r}
#| version: solution

grant_search <- nsf_terminations |> 
  filter(str_starts(as.character(grant_number), "23"))
```

```{r}
#| version: student

# 308 grants remain

```


9. The grant was intended to span four years or more. (For example, a grant that started in the year 2021 and was expected to end in 2025 spanned 5 years.)

```{r}
#| version: solution

grant_search <- grant_search |> 
  filter(expected_end_year - start_year + 1 >= 4)
```

```{r}
#| version: student

# 162 grants remain.

```


10. The grant we are looking for involves indigenous communities, so the word "indigenous" appears in the abstract. Be careful to make this search not case-sensitive!

```{r}
#| version: solution

grant_search <- grant_search |> 
   filter(str_detect(str_to_lower(abstract), "indigenous"))
```

```{r}
#| version: student

# 29 grants remain.

```

11. The title of the grant isn't tooooo long - it is  not more than 120 characters.

```{r}
#| version: solution

grant_search <- grant_search |> 
  filter(str_length(project_title) <= 120)
```

```{r}
#| version: student

# 21 grants remain.

```

12. The start date of the grant was a Friday.

```{r}
#| version: solution

grant_search <- grant_search |> 
  filter(wday(nsf_startdate, label = T) == "Fri")
```

```{r}
#| version: student

# 6 grants remain.

```


13. Finally, the word "diversity" does appear in the abstract, but even so this grant was not on Ted Cruz's infamous list.

```{r}
#| version: solution

grant_search <- grant_search |> 
  filter(diversity > 0,
         !in_cruz_list) 


grant_search |> 
  select(grant_number, project_title,
         org_name:org_state,
         usaspending_obligated) |> 
  kable()
```

```{r}
#| version: student

# 1 grant remains.

```

14. Take a look at this grant -- is our data out of date and it was actually reinstated?
(Hint: you can try searching the title or NSF grant number in a search engine, or you can look for it on the updated [Grant Witness page](https://grant-witness.us/nsf-data.html).)


:::{.version-student}

*Answer in markdown here.*

:::

:::{.version-solution}

*According to Grant Witness, as of July 2026, "Kaulele (To Take Flight) - Creating a Native Hawaiian Pacific Islander Indigenous-led Design Framework for STEM Exhibits" is still terminated, even though it was supposed to support that work through February 2029. *

:::


:::{.version-instructions}

:::callout-note
## Canvas Submission

Submit the amount of money that was still obligated to the grantees when the grant was terminated and the organizations city and state as they appear in your cleaned data. 

:::


:::


:::{.version-solution}

:::callout-note
## Canvas Submission

Submit the amount of money that was still obligated to the grantees when the grant was terminated and the organizations city and state as they appear in your cleaned data. 


\$1,440,795 was obligated to the "Kaulele (To Take Flight) - Creating a Native Hawaiian Pacific Islander Indigenous-led Design Framework for STEM Exhibits" project in Kapolei, HI.
:::
:::
